---
title: "Coming from Stata"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Coming from Stata}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
has_flextable <- requireNamespace("flextable", quietly = TRUE)
```

The development interfaces target the pinned Stata `tabtools` 2.5.1 suite
(commit `712044f8`): descriptive, regression, cell, rate, binary-outcome,
categorical, correlation, survival, effect and composite workflows.
Table structure and formatting follow the native source within the explicit
statistical, estimator and numerical boundaries below. `?tabtools` is the
command catalog; R inputs replace active Stata datasets, collection state and
local/global result macros.

This vignette translates Stata calls into R, option by option, and ends
with every place where R deliberately differs. For a whole analysis
translated, see [The Stata demo in R](#the-stata-demo-in-r).

## Three rules

1. **Option names are argument names.** `nopvalue`, `pdp`, `borderstyle`,
   `boldp`, `headershade`, and the rest keep their Stata names. A switch
   becomes `TRUE`, `opt(x)` becomes `opt = x`, and `opt("A \ B")` becomes
   either `opt = c("A", "B")` or the same Stata string.
2. **Table-building calls return a `tt_table`** (Stata's `r()`
   results are in `$stored`). `tabcell()` instead returns `tt_cell` text with
   numerical provenance; `tt_fitcount()` returns a frozen count record. `xlsx =`, `csv =` and `markdown =` also
   write it, as the Stata options do, and the table is then returned
   invisibly: assign it (`t2 <- regtab(fit, xlsx = "t2.xlsx")`) to print
   it. The object can be written later with `tt_write_xlsx()`, `tt_write_csv()` and
   `tt_write_markdown()`, or converted with `flextable::as_flextable()` and
   `tt_as_gt()`.
3. **`regtab()` takes fitted models.** Stata's `regtab` reads the active
   `collect`; R passes the models: `regtab(fit1, fit2)`.

## A first translation

A small simulated cohort stands in for a Stata dataset. Variable labels are
the `"label"` attribute (what `haven::read_dta()` sets); factors hold
categorical variables. For a `.dta` file, convert its value-labelled
predictors with `tt_as_factor(d, vars = c("drug", ...))`, which keeps the
Stata codes; leave outcomes and event indicators numeric (a 0/1 `died`
labelled No/Yes turned into a factor makes `coxph()` fit a multi-state
model).

```{r data}
library(tabtools)
set.seed(20260925)
n <- 400
cohort <- data.frame(
  treated   = factor(sample(c("SSRI", "SNRI"), n, TRUE, c(0.6, 0.4)),
                     levels = c("SSRI", "SNRI")),
  index_age = round(rnorm(n, 58, 13), 1),
  female    = rbinom(n, 1, 0.6),
  education = factor(sample(c("Primary", "Secondary", "Tertiary"), n, TRUE),
                     levels = c("Primary", "Secondary", "Tertiary")),
  diabetes  = rbinom(n, 1, 0.3)
)
cohort$cv_event <- rbinom(n, 1, plogis(-2 + 0.03 * (cohort$index_age - 58) +
                                         0.4 * cohort$diabetes))
labs <- c(treated = "Treatment", index_age = "Age at cohort entry (years)",
          female = "Female sex", education = "Education level",
          diabetes = "Diabetes", cv_event = "Cardiovascular event")
for (v in names(labs)) attr(cohort[[v]], "label") <- labs[[v]]
```

```stata
table1_tc, by(treated) vars(index_age contn %5.1f \ female bin \ education cat \ diabetes bin) ///
    smd xlsx("t1.xlsx") sheet("Table 1") title("Table 1") borderstyle(academic)
```

```{r table1}
xlsx <- tempfile(fileext = ".xlsx")
t1 <- table1_tc(cohort, by = "treated",
                vars = c(index_age = "contn %5.1f", female = "bin",
                         education = "cat", diabetes = "bin"),
                smd = TRUE, xlsx = xlsx, sheet = "Table 1",
                title = "Table 1", borderstyle = "academic")
t1
```

The `vars` specification can also be written exactly as in Stata:
`vars = "index_age contn %5.1f \\ female bin \\ education cat \\ diabetes bin"`.
A plain character vector (`vars = c("index_age", "female")`) auto-types each
variable, like a Stata varlist without `vars()`.

```stata
collect clear
collect: logit cv_event i.treated index_age female i.education diabetes
regtab, coef("OR") noint xlsx("t2.xlsx") sheet("Logistic")
```

```{r regtab}
fit <- glm(cv_event ~ treated + index_age + female + education + diabetes,
           family = binomial, data = cohort)
t2 <- regtab(fit, coef = "OR", nointercept = TRUE, xlsx = xlsx, sheet = "Logistic")
t2
```

The same object goes to Word through flextable, with the workbook's house
style (the model label merged over its columns, italic Reference cells
spanning the estimate, CI and p-value, the borders):

```{r flextable, eval = has_flextable}
flextable::as_flextable(t2)
```

## The Stata demo in R

The source repository contains an R twin of the pinned Stata demo, arranged
under corresponding command headings. Existing translated workflows remain;
the new cell/rate/outcome and analytical sections are being enrolled separately.
A command's presence in the catalog does not claim that its future demo section
has passed. The demo,
`qa/demo/demo_tabtools.R` in the GitHub repository `tpcopeland/R-tabtools`,
is not installed with the package, because its data include Stata's
example datasets `auto` and `union`, which the package does not
redistribute. From the root of a clone, with tabtools installed:

```{r, eval = FALSE}
demo <- new.env()                 # its objects stay out of your workspace
demo$out_dir <- "tabtools_demo"   # optional; the default is a folder in tempdir()
source("qa/demo/demo_tabtools.R", local = demo)
```

## `table1_tc` option by option

`desctab()` is an alias of `table1_tc()`, as `desctab` is in Stata.

| Stata | R | Notes |
|---|---|---|
| `varlist` | `vars = c("a", "b")` | auto-typed; `vars = NULL` (default) uses every column, as Stata does with no varlist |
| `vars(a contn %5.1f \ b cat)` | `vars = c(a = "contn %5.1f", b = "cat")` or the Stata string | types `auto`, `contn`, `contln`, `conts`, `cat`, `cate`, `bin`, `bine` |
| `if`/`in` | subset the data first | |
| `[fweight=w]` | `fweight = "w"` | |
| `wt(w)` | `wt = "w"` | |
| `by(g)` | `by = "g"` | |
| `total(before)`, `total(after)` | `total = "before"`, `total = "after"` | |
| `missing` | `missing = TRUE` | |
| `test`, `statistic` | `test = TRUE`, `statistic = TRUE` | R test names; see below |
| `headerperc` | `headerperc = TRUE` | |
| `smd` | `smd = TRUE` | |
| `nopvalue` | `nopvalue = TRUE` | |
| `format(%fmt)`, `percformat()`, `nformat()` | `format =`, `percformat =`, `nformat =` | Stata format strings, e.g. `"%5.1f"`, `"%12.0fc"` |
| `varlabplus` | `varlabplus = TRUE` | |
| `iqrmiddle()`, `sdleft()`, `sdright()`, `gsdleft()`, `gsdright()`, `percsign()` | same names | same defaults (`"±"`, `" (×/"`, ...) |
| `spacelowpercent`, `extraspace`, `percent`, `percent_n`, `slashN`, `catrowperc` | same names, `TRUE` | |
| `pdp(#)`, `highpdp(#)` | `pdp =`, `highpdp =` | |
| `missingsummary` | `missingsummary = TRUE` | |
| `smallcells(#)` / `smallcells(#, primary)` | `smallcells = 5, smallcells_mode = "strict"` / `"primary"` | strict is default; publication/raw boundaries below |
| `nosmallcells` | `nosmallcells = TRUE` | opt out of inherited protection |
| `cellreplace()` | `cellreplace = list(list(row = "Age", column = "Treated", text = "literal"))` | exact trimmed labels; integer value-column index excludes the label; protected targets refuse |
| `smdtype()`, `smdpair()` | same names | pair, population standardized bias or max-pair shared-scale SMD; explicit values/labels selection |
| `nosmdhighlight` | `smdthreshold = -1` | equivalent; Markdown has no SMD highlighting |
| `wtcompare`, `wtn` | `wtcompare = TRUE`, `wtn = TRUE` | |
| `smdthreshold(#)` | `smdthreshold =` | `-1` disables |
| `xlsx()`, `excel()` | `xlsx =`, `excel =` | `excel` wins when both are given, as in Stata |
| `sheet()`, `title()`, `footnote()`, `open` | same names | `sheet` may resolve an opt-in session workbook; `open` needs its resolved workbook; `title` and `footnote` need no file |
| `borderstyle()`, `font()`, `fontsize()` | same names | `borderstyle` needs no file (see "Deliberate differences") |
| `boldp(#)`, `highlight(#)`, `zebra`, `headershade`, `headercolor()`, `zebracolor()` | same names | colours as Stata names, `"R G B"` or hex |
| `csv()`, `markdown()`, `mdappend` | same names | |
| `frame(name)`, `clear` | `as.data.frame(tab)` | the table is the return value |
| `dots` | `dots = TRUE` | a progress message |
| (none) | `labels = c(a = "Age")` | override variable labels |
| (none) | `test_args = list(t.test = list(var.equal = TRUE))` | arguments for R's tests |

## `regtab`

### Which R model stands for which Stata command

The effect label (OR, HR, IRR, ...) and exponentiation follow the model, as
Stata's `regtab` decides them from the collected command.

| Stata | R | Estimate |
|---|---|---|
| `regress` | `lm()` | Coef. |
| `logit`, `logistic` | `glm(family = binomial)` | OR |
| `probit`, `cloglog` | `glm(binomial(link = "probit"))`, `"cloglog"` | Coef. |
| `poisson` | `glm(family = poisson)` | IRR |
| `glm, family() link()` | `glm()` with that family and link | by family and link |
| `nbreg` | `MASS::glm.nb()` | IRR |
| `stcox` | `survival::coxph()` (`ties = "breslow"` for Stata's default) | HR |
| `clogit` (with or without `or`) | `survival::clogit()` (`method = "exact"`, the default) | OR, as Stata since 2.1.14 |
| `streg`, `time` metric | `survival::survreg()` (Weibull, lognormal, loglogistic, exponential) | TR: `exp()` of the log-time coefficient, as Stata |
| `intreg` / `tobit` | `survreg(dist = "gaussian")` / `AER::tobit()` | Coef. |
| `stcrreg` | `cmprsk::crr()` (exact), or `coxph()` on `survival::finegray()` data | SHR |
| `ologit` | `MASS::polr()`, `ordinal::clm()` (logit link) | OR |
| `oprobit` | `MASS::polr(method = "probit")`, `ordinal::clm(link = "probit")` | Coef. |
| `mlogit` | `nnet::multinom()` | RRR |
| `zip`, `zinb` | `pscl::zeroinfl()` | Coef. |
| `churdle` layout | `pscl::hurdle()` | Coef. |
| `mixed` | `lme4::lmer(REML = FALSE)`, `nlme::lme(method = "ML")`, `glmmTMB::glmmTMB()` | Coef. |
| `melogit`, `mepoisson`, `meprobit`, `mecloglog`, `menbreg`, `meglm` | `lme4::glmer()`, `lme4::glmer.nb()`, `glmmTMB::glmmTMB()` | by family and link |
| `xtgee` | `geepack::geeglm()` | by family and link, as `glm`: OR for binomial/logit, IRR for poisson/log, else Coef. |
| `glm [pw], vce(cluster id)` read of a GEE | `geeglm(corstr = "independence")` with `gee_as = "glm"` | OR, IRR, ... by family and link |
| `svy: <cmd>` | `survey::svyglm()` | as the plain command: OR for a (quasi)binomial logit, IRR for a (quasi)poisson log; t with the design degrees of freedom |
| `teffects ipw` | `WeightIt::glm_weightit()` | Coef. (the ATE for `y ~ treat` with an identity link), M-estimation variance |
| `mi estimate: <cmd>` | a `mice` `mira` (`with(imp, glm(...))`), or `tt_mi(list_of_fits)`, of `lm`, `glm`, `coxph` or `crr` fits | as the plain command: Rubin's rules with `mi estimate`'s degrees of freedom |
| anything else | a `broom.helpers::tidy_plus_plus()`-shaped data frame | from its attributes |

Not supported: mixed-effects Cox models (`coxme`), which have no Stata
equivalent (Stata's `mestreg` is a parametric survival model, and
`stcox, shared()` is a gamma frailty model), and subclasses of the classes
above (`mgcv::gam`, frailty `coxph`) until they are verified against a
Stata command; pass those as a data frame. An `lmerTest` fit is shown as
`as(fit, "lmerMod")`, with a note (Stata `mixed`'s z statistics, not
Satterthwaite t tests). A pooled `mice` result (`mipo`) is refused: pass
the `mira`, which regtab pools itself, since `mice::pool()` applies the
small-sample degrees of freedom to every model and `mi estimate` does not.
Stata's `bootstrap:` and `jackknife:` prefixes have no R counterpart in
regtab; pass bootstrap estimates and intervals as a data frame.

**Multiple imputation.** `regtab(with(imp, glm(y ~ x, binomial)))` is
`collect: mi estimate: logit y x` followed by `regtab`: the imputations
are pooled by Rubin's rules, with Barnard and Rubin's small-sample degrees
of freedom when the completed-data model has finite residual degrees of
freedom (`regress`, `lm`) and Rubin's large-sample degrees of freedom
otherwise (`logit`, `stcox`), as `mi estimate` does; `vce = "robust"` is
`mi estimate: logit ..., vce(robust)`. `stats = c("mi_m", "fmi")` add the
Imputations and Largest FMI rows; the log-likelihood, AIC and BIC are
blank. Stata's `mi estimate` reports coefficients unless it is itself given
`or` (`mi estimate, or: logit`), and Stata's regtab shows odds ratios either
way; R has no such prefix and shows them too. See `?tt_mi`.

### Options

| Stata | R | Notes |
|---|---|---|
| `collect: <model>` | `regtab(fit1, fit2, ...)` or `regtab(list(...))` | a named list names the models |
| `models("A \ B")` | `models = c("A", "B")` or `"A \\ B"` | |
| `coef()` | `coef =` | header only; exponentiation follows the model (`coef = "AF"` over a `survreg()` fit relabels the TR column) |
| `sep()` | `sep =` | |
| `title()`, `footnote()` | same names | |
| `nointercept`, `keepintercept` | `nointercept = TRUE`, `keepintercept = TRUE` | intercept dropped by default when every model is on a ratio scale |
| `noreeffects`, `relabel` | same names | mixed models |
| `stats(n aic bic)` | `stats = c("n", "aic", "bic")` | also `events`, `groups`, `mi_m`, `qic`, `ll`, `icc`, `r2`, `r2_a`, `rmse`, `F`, `fmi`, always in Stata's row order; R-only `vce` (a "Standard errors" row) and `stat_fun =` add rows of your own |
| `digits(#)` | `digits =` | |
| `level(#)` | `level = 0.95` or `95` | Stata checks the collection's level; R fits carry none, so `level` sets the intervals |
| `keep()`, `drop()`, `labelmatch` | same names | Stata-style keys: `age`, `2.education`, `1.foreign#3.rep78` |
| `dimnonsig`, `factorlabel` | same names | analytical state determines references, never matching printed text |
| `tabtools fitcount` | `tt_fitcount(fit, events = "event", data = d, terms = TRUE)` | capture at fit time, pass as `fitcounts` |
| `mincount`, `notestlabel`, `absentlabel`, `cnslabel` | same names | explicit fit-owned states and reporting masks |
| `reftop`, `cellnote` | same names | exact placement and literal overrides; publication inference invalidated |
| `addrow ..., after()` | structured `addrow` with an exact anchor | factor anchors select the end of the block |
| `transpose`, `collabels`, `addcol` | same names | models become rows; explicit raw/equation term keys |
| `cformat`, `cilabel`, `plabel`, `statlabels` | same names | formats/custom headers; explicit digits conflict |
| rich `stats()` | structured `stats` records with `stat_values` | named scalars, formats, pairs/text, `mincell` and linked `maskwith` |
| `refcat()`, `omitlabel()`, `emptylabel()` | same names | |
| `cutlabels()` | `cutlabels =` | ordered-model cutpoints |
| `compact`, `nopvalue`, `stars`, `starslevels()` | same names | |
| `addrow("P trend" 0.03 \ ...)` | `addrow = list("P trend" = 0.03)` or the Stata string | |
| `pdp()`, `highpdp()`, `cdisc`, `labelwidth()` | same names | |
| `xlsx()`, `sheet()`, `open` | same names | regtab has no `excel()` synonym in Stata either; `excel =` is refused with a pointer to `xlsx =` |
| `borderstyle()`, `font()`, `fontsize()`, `boldp()`, `highlight()`, `zebra`, `headershade`, `headercolor()`, `zebracolor()` | same names | |
| `csv()`, `markdown()`, `mdappend` | same names | |
| `frame(name)` | `tt_flat(tab)` | keyed regression/effect rows with validated attributes; use `keyed = FALSE` for publication-only families |
| `eplotframe(name)` | `as_forest_data(tab)` | for a `comptab()` table, `model`, `model_label` and `label` are the composite's as displayed (Stata copies the source table's); the source index is in `source_model` |
| `fvgen` + `regtab` | `interactions = "fvgen"` (default), `xsymbol =`, `vsref =` | a native R interaction reproduces the fvgen-flattened table |
| native factor interactions | `interactions = "native"` | raw `1.foreign#3.rep78` keys |
| `vce(robust)`, `[pw = w]`, `vce(cluster id)` | `vce = c("stata", "model", "robust", "cluster")`, `cluster = ~id` | `"stata"`: the variance Stata reports by default; `"model"`: the fit's own `vcov()`; `"robust"`: Stata's `vce(robust)`/`[pweight]`; `"cluster"`: `vce(cluster id)`. One value, or one per model (a list or a vector as long as the models); `cluster` alone implies `"cluster"`. See [Weighted and robust variance](#weighted) |
| `xtgee` vs `glm [pw], vce(cluster)` | `gee_as = c("xtgee", "glm")` | how a `geeglm` fit is read |
| (none) | `vce_note = TRUE` | a footnote sentence naming each model's non-default variance |
| (none) | `ci_method = "profile"` | profile intervals instead of Stata's Wald |

## Output and stored results

| Stata | R |
|---|---|
| `r(methods)`, `r(Dapa)`, `r(table)`, `r(N_rows)`, ... | `tab$stored$methods`, `tab$stored$Dapa`, `tab$stored$table`, ... |
| `tabtools set font Calibri` | `tabtools_options(font = "Calibri")` |
| `tabtools set ..., permanent` / `tabtools use` | `tabtools_options(..., persist = TRUE)` (reloaded when tabtools is loaded) |
| (Excel only) | Word: `flextable::as_flextable(tab)`, then `flextable::save_as_docx()` or officer |
| HTML, Quarto | `tt_as_gt(tab)` or `flextable::as_flextable(tab)` |

```{r stored}
t2$stored$methods
head(as_forest_data(t2))
```

## Composite, rate and effect tables

Rate and model tables retain explicit identities and validated numerical
companions. `allmodels`, keyed categorical placement and `modelonly` extend
the rate scaffold. A publication flat frame is an editing/export product,
not a new model or count source.

| Stata | R | Notes |
|---|---|---|
| `puttab` (varlist, `frame()`, `matrix()`) | `puttab(x)`, `x` a data frame, matrix or `tt_table` | `subset =` for `if`/`in`; no file is required |
| `stacktab using file.xlsx, blocks(...)` | `stacktab(blocks, xlsx = "file.xlsx")` | `tt_table` blocks need no workbook |
| `strate` | `tt_rates()` | log-normal intervals, as `strate` |
| `stratetab, using(...)` | `stratetab(x)`, blocks from `tt_rates()`, strate-shaped data frames or `.dta` files | the level comes from the blocks |
| `effecttab` after `teffects` or `margins` | `effecttab()` on marginaleffects results, data frames or matrices | see [Weighted analyses](weighted-analyses.html) |
| `comptab framelist, rows()` | `comptab(list(m1, m2), rows = list(1, 3:5))` | `regtab()`/`effecttab()` tables for frames; `rownames` patterns match case-insensitive parts of labels (`"treated"` also matches `Untreated`), and the R-only `rownames_exact = TRUE` matches whole labels; see [Regression tables](regression-tables.html#table-2) |
| `hrcomptab rateframe, modelframes() rows()` (or `comptab ..., rateframe()`) | `hrcomptab(rates, list(m1, m2), rows = list(1, 3:5))` | a `coxph()` fit's outcome identity is its event variable, so `outcomemap` is often unnecessary |
| `survtab`, `crosstab`, `corrtab` | `survtab()`, `crosstab()`, `corrtab()` | See the analytical recipes below for survival, categorical association and correlations. |

## New analytical and cell workflows

```{r new-cell-and-categorical}
tabcell("np", n = c(2, 5), d = c(10, 10), ci = "exact")
categorical <- data.frame(event = c(0, 1, 0, 1), group = c(0, 0, 1, 1),
                          frequency = c(40, 10, 20, 30))
crosstab(categorical, "event", "group", weights = "frequency",
         or = TRUE, rr = TRUE, rd = TRUE)
corrtab(mtcars, vars = c("mpg", "wt", "hp"), spearman = TRUE)
```

`tabcell` covers `est`, `p`, `n`, `np`, `enp`, `iqr` and `rate`, with supplied
values, supported fitted coefficients, matrix rows or explicit contrasts.
It preserves supplied inference and redacts masked percentage/rate companions.
Crosstab's second numerical row and column define event/exposure direction;
labels never reverse it. Expected counts below five choose exact Fisher,
otherwise uncorrected Pearson. The sample OR is ad/bc. With positive cells,
its equal-tail Fisher limits follow Stata 17 `cc`: Cornfield starting values
and secant iteration stopping at a 0.5% successive-odds-ratio difference.
These returned limits can differ from `fisher.test()`'s root tolerance;
zero count cells select the native unadjusted Cornfield interval. Neither
interval changes the sample OR to the conditional-MLE point estimate.

```{r new-rates-outcomes}
cohort$years <- rep(1, nrow(cohort))
cohort$exposed <- as.integer(cohort$treated == "SNRI")
ratetab(cohort, by = "treated", events = "cv_event", exposure = "years")
outtab(cohort, outcomes = "cv_event", exposure = "exposed",
       models = list(~1, ~ index_age + female), estimator = "modified_poisson")
```

`ratetab` keeps separate sections for each grouping variable and a common
complete outcome/exposure sample. Its default is exact Poisson, distinct from
legacy `tt_rates`/`stratetab` log-normal intervals. Zero events with positive
time have exact bounds under every method; no-time cells are empty, and events
without time refuse. Cluster intervals use full fitted-cluster score sums and
G/(G-1): one event-contributing cluster can have a valid interval if other
clusters contribute exposure and score variance is positive. Supplying an ID
does not create an interval when there is only one fitted cluster or zero variance.

`outtab` distinguishes outcome-specific crude counts from each adjusted fit's
complete-case sample. The default is modified Poisson with robust variance;
`logit` and `poisson` are explicit alternatives. Built-in GLMs use
`glm.control(epsilon = 1e-12, maxit = 100)` unless `estimator_args$control`
is supplied. Diagnostic text records nonconvergence, unavailable exposure
inference, fit failure or additional sample reduction; arbitrary weighted/MI
estimators are not inferred from a display label.

```{r new-survival}
followup <- data.frame(time = c(1, 2, 3, 4), died = c(1, 0, 1, 0))
survtab(followup, "time", "died", times = c(1, 3), rmst = 3,
        median = TRUE, riskset = TRUE)
```

Survival is KM with Greenwood/log-log inference, independent-subject RMST and
full-follow-up log-rank inference. Delayed entry and split intervals require
explicit coherent subject evidence; IDs do not select robust IJ or recurrent-event
methods. RMST beyond any included group's support refuses; requested KM times
beyond support retain the final estimate with flags and a note. `reverse`
means 1−KM, not competing-risks incidence. At terminal zero survival, native
queried SE may forward-fill a previous SE for the requested contrast, while
mathematical event-grid and RMST variances retain their own values.

## Publication protection, identity and session policy

* **Strict and primary differ by command.** Table 1/crosstab strict protection
  redacts protected count returns and linked publication inference. Table 1
  primary retains explicitly labeled `stored$raw$table` and sample aggregates
  for analysis; its protected publication counts/percentages and linked ESS
  remain unavailable. Crosstab primary redacts protected returned counts and
  dependent percentages but retains computed tests/association estimates.
  Primary protection promises no complementary reconstruction guarantee.
* **Printed-only masks retain analytical data.** `stratetab`, `ratetab`,
  `outtab` and regression `mincount` preserve raw analytical returns under
  publication masks. `mincount` is reporting hygiene, not disclosure protection.
  Regression `stored$table`/raw row metadata are separate from
  `stored$publication_table`/publication rows and eligible forest inference.
  The native FC002 mincount return has four unsuppressed rows; R deliberately
  retains all six analytical rows, with the matching four-row publication
  matrix. Neither raw rows nor native identities are dropped to claim parity.
* **Overrides use canonical state.** Table 1 replacements class-refuse primary,
  complementary or linked-withheld targets before any mutation, an intentional
  safer difference from native unchecked replacement. `cellreplace` uses R
  records with exactly row/column/text; numeric column1 means the first value
  column, whereas character `"1"` means an exact header. Regression `cellnote`
  clears publication inference and forest eligibility while preserving raw
  analysis. A real estimate printed `1.00` is never inferred to be Reference.
* **Later-source strict count guards.** Table 1 ports the denominator and
  printed-continuous-summary guards from pinned tabtools 2.5.2–2.5.6 source
  (`4eecca4d`). Strict `slashN` withholds small denominators and denominators
  beside masked N, including a reconstructive Total; visible Missing counts
  omit percentages when their denominator is withheld. A printed continuous
  summary contributes the released nonmissing-count lower bound to strict
  feasibility. These selected repairs do not establish universal protection
  beyond supported layouts or complete equivalence to 2.5.6. Primary mode
  retains its documented limited protection, and the finite-SD R choice stays.
* **Flat frames have truthful scope.** Keyed regression/effect rows preserve raw
  equation keys, original model/state matrices and unique occurrence blocks.
  Puttable technical keys are omitted only when actually selected and omitted-key
  diagnostics are separate paragraphs in every R sink (native console only).
  Transposed regression and composite/rate/outcome/categorical/correlation/survival
  publications require plain unkeyed export (`tt_flat(x, keyed = FALSE)`).
  No artificial regression model IDs/states are assigned. Composites authenticate
  numerical companions and classify references from canonical flags, not labels;
  keyed/model-only/multiple-model-per-outcome forest requests refuse.
* **Session opt-in and explicitness.** `tabtools_options()` queries the defaults;
  use `tabtools_options()$digits`, not unnamed `tabtools_options("digits")`.
  Explicit named NULL clears a session key. Ordinary commands inherit workbook
  and Markdown only for an explicitly supplied non-NULL sheet. Their omitted
  default sheet causes no inherited write. Explicit destination NULL opts out.
  `puttab` and `stacktab` frame publication has its documented omitted-destination
  inheritance. Missing workbook for an explicit sheet warns; internal staging
  never triggers a second hidden write. Successful session Markdown writes
  replace/create first and append subsequently, with explicit append winning.
  Workbook alias priorities are command-specific: non-NULL excel wins for
  Table 1, nonempty xlsx first for cross-tabulation/correlation. Explicit
  `excel = NULL` alone opts Table 1 out of inherited workbooks.
  An explicit active `smallcells = k` defaults to strict unless the call
  explicitly selects a mode. Omitted threshold/mode inherit together;
  explicit NULL/0 or `nosmallcells` disables inheritance, with the documented
  argument-conflict rules. Publication notes describe threshold/protection
  limits; analytical backup field details belong in help, not those notes.
* **Paragraphs are cross-sink.** A footnote character vector or the reserved
  single spaced-backslash token makes separate paragraphs. Unspaced/doubled
  backslashes, dollar signs, quotes and backticks remain literal data.
  R carries complete notes through text, workbook and presentation renderers;
  some native console-only diagnostics become all-sink R paragraphs.
* **Inference is selected, not generic.** Table 1 retains the R test map below.
  Corrtab recomputes pairwise tied ranks and N; the chosen Stata17 Spearman
  engine uses t(N−2), unlike the current manual's beta method or R AS89/exact
  inference. Corrtab N2 p-values are missing; for N>2 perfect positive
  Spearman p=0 while perfect negative Spearman p is unavailable. Constants
  have blank undefined diagonals and useful pair counts. Crosstab refuses
  unavailable N<=2/negative-perfect trend inference before sinks.
* **Estimator limits are explicit.** `effecttab(full)` remains the deliberate
  `regtab()` component-fit alternative, with equation-aware term identities.
  Interval-censored supported `survreg` AFT fits use TR; Gaussian interval
  regression uses Coef. Unsupported native estimator families require an
  explicit custom-frame effect scale, confidence level, inference and survival
  metric metadata; no class support is inferred from a header. MI varying-sample
  `esampvaryok` and pooled-MI mincount remain unsupported. `lmerTest` objects
  retain the documented Stata mixed-model z policy, not Satterthwaite inference.
  Ratio-scale effect SEs retain delta-method Wald limits and warn on a
  nonpositive derived lower bound. Logical Table 1 groups remain 0/1; SMD
  selectors accept TRUE/FALSE (logical or strings) as1/0.

## Deliberate differences

Everything above aims at identical output. These are the places where R
does something else on purpose, and why.

### Hypothesis tests (table1_tc)

Displayed descriptive statistics match Stata; p-values come from R's
conventional tests (only the numbers: the p-value format, `pdp`/`highpdp`,
`<0.001`, is Stata's). The Test column and the methods paragraph name the R
test, and the paragraph drops Stata's closing software sentence.

| Variable type | Groups | Stata (Test column) | R (Test column) |
|---|---|---|---|
| `contn`, `contln` | 2 | pooled-variance t via `regress` (`Ind. t test`) | `t.test()`, Welch (`Welch t test`) |
| `contn`, `contln` | >2 | `anova` (`ANOVA`) | `oneway.test()`, Welch (`Welch ANOVA`) |
| `conts` | 2 | `ranksum` (`Wilcoxon rank-sum`) | `wilcox.test()` (`Wilcoxon rank-sum`); exact for n < 50 without ties, with continuity correction |
| `conts` | >2 | `kwallis` (`Kruskal-Wallis`) | `kruskal.test()` (`Kruskal-Wallis`) |
| `cat`, `bin` | any | `tab, chi2` (`Chi-square`) | `chisq.test(correct = FALSE)` (`Pearson's chi-squared`), the same test |
| `cate`, `bine` | any | `tab, exact` (`Fisher's exact`) | `fisher.test()` (`Fisher's exact`); a table too large for the exact algorithm gets a Monte Carlo p-value under a fixed internal seed, labelled `Fisher's exact (simulated)` |

`test_args = list(t.test = list(var.equal = TRUE), oneway.test =
list(var.equal = TRUE))` gives Stata's pooled-variance tests. Stata's
`kwallis` keeps ranks in float precision, so its p-values differ from
`kruskal.test()` around the seventh significant digit.

### Input that R refuses or treats differently

- **Invalid display formats** are an error. Stata's `string()` silently
  returns `""` for a format such as `%5.1x` or `%5.5g` (decimals not
  fewer than the width), so the cell prints as `±`.
  Fixed, general and exponential formats are supported; `%ec` is refused.
  General formats requesting at least 13 significant digits can differ in
  final rounding digits. Allocation limits are widths 1–10000 and decimals
  0–400; a zero-decimal general format has width at most 402.
- **Dates** (`Date`, `POSIXt` columns) are refused: Stata counts days from
  1960 and R from 1970, so no summary could match. Convert them first.
- **Non-finite values**: `Inf`, `-Inf`, `NaN`, and values of 2^1023
  (8.99e307) or more, in any analysed column (`by` included) or in a
  weight, are refused by name. A Stata variable cannot hold them; a Stata
  weight of 1e308 is a missing value that silently drops its record.
- **Centered SD overflow**: by explicit user choice, R retains a finite SD
  where the pinned native square of the weighted centered-deviation sum
  overflows and prints missing, but its variance correction remains finite.
  Ordinary centered calculations prevent cancellation at large offsets.
  This finite-SD boundary is an intentional difference, not an assertion of
  native missing-value equivalence. Weighted rescaling and other unavailable
  sum/SMD boundaries follow their documented command contracts.
- **SMDs under extreme weights**: for accepted weights near `8e307`, R uses
  bounded rescaling to retain finite SMDs where Stata 2.5.6 returns missing
  values, separately from the finite SD choice.
- **Tiny weights and huge values** follow Stata tabtools (since 2.1.14):
  a weight total below 1 is multiplied by a power of two before the
  weighted quartiles (1..10 under weights of 1e-12 prints `6 (3, 8)`, as
  for any equal weights), and automatic typing divides the
  values by a power of two before the moments and the Shapiro-Wilk test,
  so values near 1e100 are typed by their shape.
- **Output files** are checked before anything is written: `csv` must name
  a `.csv` file and two arguments naming the same file are refused, as in
  Stata tabtools (since 2.1.12); R also refuses a missing or read-only directory
  there. An existing Markdown file is replaced unless `mdappend`, as in
  Stata.
- **Paired tests and `mu`** (R-only `test_args`): `paired = TRUE` for
  `t.test` or `wilcox.test` pairs the i-th record of each of two groups in
  row order and is labelled `Paired t test` / `Wilcoxon signed-rank`;
  unequal groups, more than two groups, and (under `fweight`) pair members
  with different frequencies are errors. A nonzero `mu` is refused.
- **Stricter checks on R objects**: `labels =` overrides must be named,
  non-missing strings; duplicate column names in `data` are refused for
  the analysed columns; `validate_tt_table()` refuses missing roles, row
  types, `NA` text, and fractional or missing header spans.
- **Level order**: numeric and labelled variables follow the value codes,
  and R factors keep their own level order (Stata has no such type). In
  `table1_tc`, character variables sort bytewise (Stata's `encode`). In
  `regtab`, a character predictor takes the level order of the fit, which
  R sets from the locale's collation: under `en_US.UTF-8`, `c("a", "A",
  "b", "B")` sorts `a A b B` where Stata's `encode` gives `A B a b`, so the
  **reference level itself** can differ from Stata's. Convert the column
  before fitting with `factor(x, levels = sort(unique(x), method =
  "radix"))` to get Stata's order.
- `r(table)` row names cut at 32 bytes drop a partial UTF-8 character
  instead of keeping a stray byte.
- `regtab(models =)` with more labels than models warns (Stata errors).
- **`title`, `footnote` and `borderstyle` without a file**: Stata's
  `table1_tc` refuses `title()` without `excel()` or `markdown()` and
  `borderstyle()` without `excel()`, because a Stata table has nowhere else
  to go. In R the table goes on to `flextable::as_flextable()`, `tt_as_gt()` or
  a later `tt_write_xlsx()`, so `table1_tc()` accepts them without a file,
  as `regtab()` already did. `sheet`, `open` and `mdappend` still need their
  resolved file.

### Rate tables (stratetab)

- **Default exposure and outcome labels**: Stata's `stratetab` labels the
  blocks `Exposure 1`, `Exposure 2`, ... and the outcomes `Outcome 1`, ...
  unless `explabels()` and `outlabels()` are given, because a `strate` file
  records no variable names. A `tt_rates()` result from time and event data
  knows its grouping column and event variable, so when every block is
  one, `stratetab()` names each exposure by the grouping column's `"label"`
  attribute (else its name) and each outcome by the event column's label
  (else its name); `tt_rates()` keeps the grouping column's label for this.
  Strate-shaped data frames and `.dta` files, a mix of the two kinds of
  block, a block without a grouping column, and names that would repeat
  (the same `by` in every exposure, the same event in every outcome) keep
  Stata's defaults. Given `explabels` and `outlabels` always win. On the
  same route the default outcome identities are the event variables' names
  (Stata: the outcome labels), which are a `coxph()` model's outcome
  identities, so `hrcomptab()` pairs rates and models without `outcomemap`.
- **Scale check**: blocks from `tt_rates(per = k)`, `k` not 1, give a
  warning when a numeric `unitlabel` matches neither `k * ratescale` nor
  `k * ratescale / pyscale` (`per = 1000` with `ratescale = 1000` shows
  rates per million under "Per 1,000 PY"). Stata cannot check this: a
  `strate` file does not record its `per()`.

### Composite and effect tables (comptab, hrcomptab, effecttab)

- **Hazard-ratio check**: `hrcomptab()` refuses an `effecttab()` model of
  differences or predictions even when its `effect` header says HR (Stata
  trusts the header), and an indicator row placed on a block whose
  categories read Yes/No, 1/0 or TRUE/FALSE, where Stata's rule (an
  indicator is the second category) would put the treated effect on "No".
- **`effecttab(estimand =)`** (R only) names the effect a result estimates
  (ATE, ATT, ATC) when the result cannot say, for example an
  `avg_comparisons()` over the treated.
- **Data-frame p-values**: a p-value derived from `std.error` tests the
  frame's `null.value` column (default 0); a ratio header without one is
  refused.

### Output files

- **CSV formula cells**: `tt_write_csv()` and the commands' `csv =` write a
  leading `'` before a cell a spreadsheet would run as a formula
  (`=HYPERLINK(...)`, `+cmd|...`, `@SUM(...)`); Stata's `export delimited`
  writes it as it is. Numbers, intervals and labels such as
  `+ Comorbidities` are written unchanged.

### Models (regtab)

- **Standard errors** are the ones Stata reports (`vce = "stata"`): the
  observed information for non-canonical links and `glm.nb`, the normal
  distribution for every `glm` family, Stata's `M/(M - 1)` factor on robust
  `coxph` and `survreg` fits. `vce = "model"` uses the fit's own `vcov()` and R's
  `summary()` distribution instead; `vce = "robust"` and `"cluster"` give
  Stata's `vce(robust)`/`[pweight]` and `vce(cluster id)` variances (see
  [Weighted and robust variance](#weighted)).
- **Perfect prediction**: Stata's `logit` drops observations a covariate
  predicts perfectly; R's `glm()` keeps them. regtab warns and prints the
  coefficient with a blank interval.
- **Convergence**: R's `glm()` stops earlier than Stata for non-canonical
  links; refit with `glm.control(epsilon = 1e-12)` when a last digit
  differs. `polr`, `multinom` and `zeroinfl` fits also stop slightly short
  of the maximum.
- **Weights in `glm()`** are analytic weights under the default
  `vce = "stata"` (Stata's `[aweight]`): `Observations` counts rows with a
  positive weight, not their sum. For inverse-probability weights
  (Stata's `[pw]`) use `vce = "robust"`; regtab warns when non-integer,
  non-constant weights meet the default, but never switches by itself.
  The log-likelihood, AIC and BIC are those of Stata's `glm [aweight]`
  (weights rescaled to mean 1), and a Gamma or inverse Gaussian `glm`
  reports Stata's log-likelihood at scale 1, not R's `logLik()`.
- **Weights in `polr()`, `clm()`, `multinom()`, `zeroinfl()`, `hurdle()`
  and `survreg()`** are frequency weights, Stata's `[fweight]`:
  `Observations` (and BIC) use the sum of the weights, and the standard
  errors are those of an expanded data set. Inverse-probability weights
  passed there give a wrong analysis: an inflated N and standard errors that
  are too small. regtab warns about non-integer weights on these fits and
  refuses `vce = "robust"`/`"cluster"` for them (it has no robust variance
  there); refit `survreg()` with `robust = TRUE` for Stata's robust `streg`.
- **Ordered factors** and non-treatment contrasts are refused: a Reference
  row claims a contrast against that level. Use an unordered factor.
- **`stcrreg`**: `cmprsk::crr()` reproduces Stata exactly; `coxph()` on
  `finegray()` data approximates it (its variance omits the
  censoring-distribution term). `crr` fits carry no labels, so their rows
  show covariate names.
- **lme4 and glmmTMB**: for probit and cloglog links and for `glmer.nb()`,
  lme4's Laplace approximation is not Stata's, so the last digit can differ
  (regtab says so once per session); `glmmTMB()` reproduces Stata's Laplace
  fits exactly. `glmer(nAGQ = 7)` corresponds to Stata's default adaptive
  quadrature; `nAGQ = 0` is refused. lme4's default optimizer can stop short
  on random slopes: `lmerControl(optimizer = "bobyqa")` reaches Stata's
  optimum.
- **`nlme::lme()`**: regtab shows the fit's `varFix` standard errors with z
  statistics, as `mixed` does. `summary.lme()` prints different ones for
  an ML fit (times `sqrt(N / (N - p))`, with t tests), so its table and
  regtab's disagree by design. nlme's default is REML (`mixed, reml`); fit
  with `method = "ML"` for Stata's default. `correlation =` and `weights =`
  structures are refused (Stata's `residuals()` option has no rows in
  regtab's layout). On unscaled random slopes lme can stop short too:
  `lmeControl(niterEM = 200)` reaches Stata's optimum.
- **`streg, time` ancillary rows (`survreg()` with `keepintercept`)**: the
  covariate rows are time ratios, `exp()` of the log-time coefficients,
  and a kept intercept is `exp()` of the constant; the ancillary rows
  (`ln_p`, `p`, `1/p`; `lnsigma`, `sigma`; `lngamma`, `gamma`) stay on
  their own scale, as in Stata tabtools since 2.1.12.
- **`xtgee`**: `geeglm()` is shown by glm's family/link rule, as Stata
  tabtools (since 2.1.12) shows `xtgee`: odds ratios for binomial/logit and
  incidence rate ratios for poisson/log, intercept dropped, coefficients
  otherwise, with
  xtgee's model-based variance; `vce = "model"` gives geepack's sandwich,
  which is Stata's `vce(robust)` without its `G/(G - 1)` factor. QICu has
  Stata's additive constant only for fixed-scale families. A weighted
  `geeglm` is `xtgee [pw]`, robust by default. `gee_as = "glm"` reads an
  independence-correlation fit as Stata's `glm [pw], vce(cluster id)`
  instead: ratios by family and link, the sandwich times `G/(G - 1)`.
- **Methods sentence** (`r(methods)`, `tab$stored$methods`): built from
  the models, as in Stata tabtools since 2.1.12: the estimates by each model's effect
  scale (a `coef` or `cdisc` relabel does not change them), the model by
  its command, family and link, "univariable" or "multivariable" by the
  number of predictor variables. The R-only models name themselves
  (`pscl::hurdle()` count models are not Stata's `churdle`; `clm()` links
  Stata has no ordered model for).
- **Data-frame input**: the ratio scale comes from `effect_scale`,
  `stata_cmd` or broom.helpers' `exponentiate` record (exponentiated
  `broom::tidy()` output declared as coefficients is refused); on a ratio
  scale `std.error` is used only when its scale is declared
  (`attr(x, "se_scale")`: `"link"` for broom's `exponentiate = TRUE`,
  `"estimate"` for `parameters`); supplied intervals fix the
  confidence level (`attr(x, "conf.level")`, which `tidy_plus_plus()`
  sets); a `df` column gives t statistics; factor levels are keyed
  `2.treat`, so `regtab(fit, tidy_plus_plus(fit))` shares their rows.
  Malformed columns (factors, reversed bounds, p outside \[0, 1\], ...) are
  refused. See `?regtab`, "Data-frame input".
- **Upstream Stata quirks not reproduced**: with a one-level mixed model
  beside a two-level one, Stata loses the second model's outer variance and
  labels its inner one `var(_cons)`; R shows both.
- **Covariates named like intercepts or ancillary parameters** (`p`,
  `alpha`, `constant`, `cut1`, ...) are ordinary covariates: kept by
  `nointercept` and exponentiated with the others, as in Stata tabtools since 2.1.12,
  which decides these rows from the coefficient's equation and name. A
  covariate named like an ancillary parameter of its own or another model
  (`alpha` in a Poisson model beside an `nbreg`) keeps its own row, apart
  from the parameter's, as in Stata tabtools since 2.1.13, and `keep = "alpha"` selects
  both.
- **Pseudo R-squared with an offset** is blank for `ologit`, `oprobit`,
  `logit` and `probit` models, as Stata reports none, and against the
  offset-only model for `poisson`. `multinom()` offsets (Stata's `mlogit`
  refuses `offset()`) enter the standard errors and the constant-only
  model.
- **Data changed after fitting**: `coxph()`, `survreg()`, `multinom()`
  and `model = FALSE` fits keep no copy of their data, so regtab reads
  the data again and refuses the fit when they no longer reproduce it;
  fit with `model = TRUE` to tabulate after changing the data. A
  `coxph()` with `tt()` terms (survival allows no `model = TRUE` there) is
  checked on its observations and events, and gets the variances survival
  stores (`vce = "stata"`/`"model"`, a fit with `cluster =` or
  `robust = TRUE`).
- **Fine-Gray `coxph()` fits** are recognised by their weights, their
  `Surv(start, stop, status)` response and `finegray()`'s mark on the
  data, not by the `fg*` column names (`finegray = TRUE` declares one);
  keep the subject id in the `finegray()` formula and fit with
  `cluster = id` (or `id = id`) for delayed entry, counting-process input,
  or expanded data re-sorted with new row names; without it regtab reads
  the subjects only from right-censored `finegray()` output.
- **Pseudo R-squared of `cloglog`** (and of a `glm` whose link is not
  logit/probit or, for Poisson, log): none, as Stata's `cloglog` and
  `glm` report none.
- **`xtgee, scale(#)`** is `geeglm(scale.fix = TRUE, scale.value =
  rep(#, nrow(data)), data = data)` with an independence working
  correlation. geepack stores only the `scale.value` expression, so it may
  refer only to the data (a scale held in another variable is refused).
  With a working correlation other than independence, geepack's
  fixed-scale fit estimates the correlation differently from xtgee (a note
  says so). A `geeglm()` fit on data not sorted by the id is refused
  (geepack would split an id into several clusters).
- `keep = "i.a##i.b"` matches nothing and errors (Stata keeps only the
  parent row).

## Weighted and robust variance (IPTW, MSM) {#weighted}

R's `glm(weights = w)` does not record what kind of weight `w` is, while
Stata's `[pw = w]` always reports a robust variance. So `vce` says it, per
model. Under the default `vce = "stata"` a weighted `glm()` gets Stata's
analytic-weight (`[aw]`) variance, which differs from Stata's `[pw]` robust
standard error (for unstabilised weights it understates it).

| Stata | R |
|---|---|
| `logit y i.treated [pw = w]` | `regtab(glm(y ~ treated, binomial, d, weights = w), vce = "robust")` |
| `regress y x [pw = w]` | `regtab(lm(y ~ x, d, weights = w), vce = "robust")` (HC1, t) |
| `logit y x [pw = w], vce(cluster id)` | `regtab(fit, vce = "cluster", cluster = ~id)` |
| `stcox x [pw = w]` | `regtab(coxph(Surv(time, died) ~ x, d, weights = w, robust = TRUE))` |
| `svy: logit y x` | `regtab(survey::svyglm(y ~ x, design, family = quasibinomial))` |
| `teffects ipw` weights from WeightIt | `regtab(WeightIt::glm_weightit(...))` |

```{r weighted}
# Stabilised IPTW weights for treatment.
ps <- glm(treated ~ index_age + female + education + diabetes,
          family = binomial, data = cohort)$fitted.values
p1 <- mean(cohort$treated == "SNRI")
cohort$sw <- ifelse(cohort$treated == "SNRI", p1 / ps, (1 - p1) / (1 - ps))

crude <- glm(cv_event ~ treated, family = binomial, data = cohort)
adjusted <- glm(cv_event ~ treated + index_age + female + education + diabetes,
                family = binomial, data = cohort)
# (binomial with non-integer weights warns about non-integer successes;
# the estimates are those of Stata's logit [pw = sw].)
iptw <- suppressWarnings(glm(cv_event ~ treated, family = binomial,
                             data = cohort, weights = sw))

# One vce per model: model-based for the unweighted fits, robust
# (Stata's [pw] HC0 x N/(N-1)) for the weighted one.
tw <- regtab(crude, adjusted, iptw, models = c("Crude", "Adjusted", "IPTW"),
             vce = list("stata", "stata", "robust"), coef = "OR",
             nointercept = TRUE, vce_note = TRUE)
tw
# vce_note: the footnote (workbook, Word, HTML) names the robust column.
tw$footnote
```

```{r msm}
# Pooled-logistic marginal structural model on person-period data, with
# the baseline weights carried forward and variance clustered on the person
# (Stata: logit y i.treated i.period [pw = sw], vce(cluster id)).
# A small person-period data set: one row per person and year.
pp <- cohort[rep(seq_len(nrow(cohort)), each = 3), c("treated", "sw")]
# Row indexing drops the variable label; restore it for the table.
attr(pp$treated, "label") <- attr(cohort$treated, "label")
pp$id <- rep(seq_len(nrow(cohort)), each = 3)
pp$period <- rep(1:3, nrow(cohort))
pp$event <- rbinom(nrow(pp), 1, 0.05)
msm <- suppressWarnings(glm(event ~ treated + factor(period), family = binomial,
                            data = pp, weights = sw))
regtab(msm, models = "MSM", vce = "cluster", cluster = ~id, keep = "treated")

# The same variance matrix, e.g. for marginaleffects (Stata's margins):
V <- tt_vcov(msm, vce = "cluster", cluster = ~id)
sqrt(diag(V))[2]
```

With `vce = "stata"`, a `glm()` or `lm()` whose weights are non-integer and
not constant keeps the analytic reading but warns that the weights look
like inverse-probability weights, rather than switching silently.
`tt_vce_types(fit)` lists the `vce` values a fit accepts.
