---
title: "Compared with gtsummary"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Compared with gtsummary}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
ok <- requireNamespace("gtsummary", quietly = TRUE) &&
  requireNamespace("gt", quietly = TRUE) &&
  requireNamespace("flextable", quietly = TRUE)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = ok)
# Inline values below are read from the tables built here.
iv <- function(x) if (ok) x else "..."
```

gtsummary is the most widely used R package for the same two tables. This
vignette builds one Table 1 and one Table 2 with each package, from the
same data and the same models, and lists where the rows differ. Each table
is shown in its package's usual HTML form: gtsummary through gt, tabtools
through flextable (the workbook's house style).

```{r data}
library(tabtools)
set.seed(20260925)
n <- 600
d <- data.frame(
  treated   = factor(sample(c("SSRI", "SNRI"), n, TRUE, c(0.6, 0.4)),
                     levels = c("SSRI", "SNRI")),
  age       = round(rnorm(n, 58, 13), 1),
  female    = rbinom(n, 1, 0.6),
  education = factor(sample(c("Primary", "Secondary", "Tertiary"), n, TRUE),
                     levels = c("Primary", "Secondary", "Tertiary")),
  diabetes  = rbinom(n, 1, 0.3)
)
d$event <- rbinom(n, 1, plogis(-2 + 0.03 * (d$age - 58) + 0.4 * d$diabetes))
labs <- c(treated = "Treatment", age = "Age at cohort entry (years)",
          female = "Female sex", education = "Education level",
          diabetes = "Diabetes", event = "Cardiovascular event")
for (v in names(labs)) attr(d[[v]], "label") <- labs[[v]]
```

## Table 1

```{r t1-gtsummary}
g1 <- gtsummary::tbl_summary(
  d[, c("treated", "age", "female", "education", "diabetes")],
  by = treated,
  statistic = list(age ~ "{mean} ({sd})"),
  digits = list(age ~ 1)
)
g1 <- gtsummary::add_p(g1)
gtsummary::as_gt(g1)
```

```{r t1-tabtools}
t1 <- table1_tc(d, by = "treated",
                vars = c(age = "contn %5.1f", female = "bin",
                         education = "cat", diabetes = "bin"))
flextable::as_flextable(t1)
```

```{r t1-values, include = FALSE}
tb1 <- gtsummary::as_tibble(g1, col_labels = FALSE)
g_cell <- function(lab, col = "stat_1") tb1[[col]][tb1$label == lab]
t_cell <- function(tab, lab, j = 2L) tab$body[[j]][trimws(tab$body[[1]]) == lab]
g_p <- paste0("`", stats::na.omit(tb1$p.value), "`", collapse = ", ")
tp <- t1$body[[ncol(t1$body)]]
t_p <- paste0("`", tp[nzchar(tp)], "`", collapse = ", ")
n1 <- sum(d$treated == "SSRI")
```

Row by row:

| | gtsummary | tabtools |
|---|---|---|
| Label column header | `Characteristic` | the descriptor, `No. (Column %) or Mean±SD`, merged over both header rows |
| Group sizes | in the group's header cell: `SSRI`, then `N = `r iv(n1)`` on a second line | a second header row: `N=`r iv(n1)`` under each group label |
| What the cells hold | a footnote on each column: `Mean (SD); n (%)` | the descriptor; no footnote |
| Continuous cell | `` `r iv(g_cell(labs[["age"]]))` `` | `` `r iv(t_cell(t1, labs[["age"]]))` `` (Stata's `sdleft`) |
| Categorical cell | `` `r iv(g_cell("Primary"))` `` | `` `r iv(t_cell(t1, "Primary"))` `` (Stata's default `percsign("")`) |
| Label row of a categorical variable | the variable label, p-value on it | the same |
| Level rows | indented when the table is rendered (four no-break spaces in gt output, cell padding in flextable, none in `as_tibble()`) | three literal spaces in the label text itself, kept in every output: console, CSV, Markdown (`&nbsp;`), Excel, Word, HTML |
| Binary variable | one row (the level 1 count) | one row (`bin`) |
| p-value, continuous | Wilcoxon rank-sum (default for continuous variables, whatever statistic is shown) | Welch t test for `contn`, Wilcoxon for `conts` |
| p-value, categorical | Pearson's chi-squared without continuity correction, or Fisher's exact test when an expected count is below 5 | Pearson's chi-squared without continuity correction for `cat`/`bin` whatever the expected counts (as Stata's `tab, chi2`); Fisher's exact only for `cate`/`bine` |
| p-values shown | `r iv(g_p)` (one decimal from 0.1 up) | `r iv(t_p)` (`pdp`/`highpdp`: two decimals from 0.10 up, up to three below, as in Stata) |

tabtools reads its variable types from `vars` (or auto-types them as Stata
does), while gtsummary picks a type and a test per variable and is
configured with `type`, `statistic` and `test`. The tests can be aligned:
`add_p(test = list(age ~ "t.test"))` in gtsummary, or `conts` in tabtools.

## Table 2

Two logistic models, crude and adjusted.

```{r models}
crude <- glm(event ~ treated, family = binomial, data = d)
adjusted <- glm(event ~ treated + age + female + education + diabetes,
                family = binomial, data = d)
```

```{r t2-gtsummary}
g2 <- gtsummary::tbl_merge(
  list(gtsummary::tbl_regression(crude, exponentiate = TRUE),
       gtsummary::tbl_regression(adjusted, exponentiate = TRUE)),
  tab_spanner = c("Crude", "Adjusted"),
  quiet = TRUE
)
gtsummary::as_gt(g2)
```

```{r t2-tabtools}
t2 <- regtab(crude, adjusted, models = c("Crude", "Adjusted"), coef = "OR")
flextable::as_flextable(t2)
```

```{r t2-values, include = FALSE}
tb2 <- gtsummary::as_tibble(g2, col_labels = FALSE)
lab_d <- labs[["diabetes"]]
g_ci <- tb2$conf.low_2[tb2$label == lab_d]
t_ci <- t_cell(t2, lab_d, 6L)
```

Row by row:

| | gtsummary | tabtools |
|---|---|---|
| Model headers | a spanner per merged table (`tab_spanner`) | each model's label merged over its OR, CI and p-value columns |
| Reference level | an em dash in the OR and CI cells | the word `Reference`, italic, in one cell merged across that model's OR, CI and p-value columns |
| Level rows | indented by the renderer | two literal spaces in the label text |
| A term absent from a model | blank cells | blank cells |
| Confidence interval | profile likelihood (`confint()` of the `glm`); Diabetes, adjusted: `` `r iv(g_ci)` `` | Wald, as Stata reports and consistent with the p-value: `` `r iv(t_ci)` ``; `ci_method = "profile"` gives profile intervals |
| Row order | each model's terms, merged | the union of all models' rows in order of first appearance, intercept last |
| Model statistics | `add_glance_table()` or `add_glance_source_note()` | `stats = c("n", "aic", "bic")` rows below the coefficients, merged across each model's columns |
| Footnote | abbreviations (`CI = Confidence Interval, OR = Odds Ratio`) | only what `footnote`/`stars` add |

## Where the difference matters

Both packages produce publication tables. The differences above are the
reason tabtools exists: it reproduces the Stata `tabtools` layout, so a
table built in R matches one built in Stata cell for cell, and the same row
structure is kept in every output (console, CSV, Markdown, Excel, Word,
HTML). gtsummary covers more summary types, more model types, and many ways
to extend and restyle a table that tabtools does not attempt.

A tabtools table can also be shown with gt (`tt_as_gt(t2)`; gtsummary's
own `as_gt()` converts gtsummary tables only, and tabtools does not mask
it), with one limitation: gt cannot merge body cells, so the Reference
text sits in the OR column instead of spanning the model's columns.
`tt_as_gtsummary(t2)` turns a tabtools table into a gtsummary table, with
tabtools' cells unchanged, for gtsummary's own tools such as `tbl_stack()`
and the `modify_*()` functions.
