---
title: "Regression tables with regtab()"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Regression tables with regtab()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
# Every example uses survival (Suggests): without it the code is shown,
# not run.
has <- function(...) all(vapply(c("survival", ...), requireNamespace, TRUE, quietly = TRUE))
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = has())
```

`regtab()` turns one or more fitted models into a regression table: one
block of estimate, confidence-interval and p-value columns per model, a label
row and indented level rows for each factor, a Reference row for each base
level, and optional model statistics below. This article covers which models
it reads and how it labels them, how rows from several models are combined,
the options that select and label rows, the standard errors, the numbers
behind the table, and what to do with a model regtab does not support.

```{r setup}
library(tabtools)
library(survival)
```

## Supported models and their effect measures

regtab reads the effect measure from the model, as Stata's `regtab` reads it
from the estimation command. Ratio measures are exponentiated, with
exponentiated confidence limits; everything else is shown on the scale of the
linear predictor, labelled `Coef.`. The intercept row is dropped by default
when every model in the table is on a ratio scale.

| Family | R fit | Header |
|---|---|---|
| Linear | `lm()` | Coef. |
| Binary | `glm(family = binomial)` (logit), `quasibinomial` | OR |
| | `glm(binomial(link = "probit"))`, `"cloglog"`, `"log"` | Coef. |
| | `survival::clogit()` (conditional logistic) | OR |
| Other GLMs | `glm()` with any other family or link (Gamma, Gaussian with a log link, ...) | Coef. |
| Count | `glm(family = poisson)` (log), `quasipoisson`, `MASS::glm.nb()` | IRR |
| | `pscl::zeroinfl()`, `pscl::hurdle()` | Coef. |
| Survival | `survival::coxph()` | HR |
| | `coxph()` on `survival::finegray()` data, `cmprsk::crr()` | SHR |
| | `survival::survreg()`: Weibull, exponential, lognormal, log-logistic (accelerated failure time) | TR |
| | `survreg(dist = "gaussian")` (Stata `intreg`), `AER::tobit()` | Coef. |
| Ordinal | `MASS::polr()`, `ordinal::clm()` with a logit link | OR |
| | the same with a probit or other link | Coef. |
| Multinomial | `nnet::multinom()` | RRR |
| Mixed | `lme4::lmer()`, `nlme::lme()` | Coef. |
| | `lme4::glmer()`, `lme4::glmer.nb()`, `glmmTMB::glmmTMB()` | by family and link, as the GLMs above |
| GEE | `geepack::geeglm()` | by family and link, as the GLMs above (Stata `xtgee`) |
| Survey | `survey::svyglm()` | as the GLMs above; t with the design degrees of freedom |
| Weighting | `WeightIt::glm_weightit()` | as the GLMs above, with the M-estimation variance |
| Anything else | a data frame shaped like `broom.helpers::tidy_plus_plus()` output | from its attributes (see [below](#data-frame-input)) |

Three consequences are worth knowing before you fit:

- A **log-binomial** model (`binomial(link = "log")`) is shown as log risk
  ratios under `Coef.`, not as risk ratios, because Stata's `regtab` does
  the same for `glm, link(log)`. For a risk ratio table, fit a Poisson
  model with robust standard errors ("modified Poisson"; Zou G, 2004,
  *American Journal of Epidemiology* 159(7), 702-706) and
  relabel the header with `coef = "RR"`; see
  [Standard errors](#standard-errors).
- `coef` changes only the header, never whether estimates are
  exponentiated.
- **GEE** fits follow Stata's `xtgee`, which Stata's `regtab` classifies
  like `glm`: odds ratios for a binomial logit, rate ratios for a Poisson log
  link, coefficients otherwise; an independence-correlation fit can instead
  be read as Stata's `glm [pw], vce(cluster id)` with `gee_as = "glm"`.

The methods sentence stored with each table (`tab$stored$methods`) names the
model and its effect measure.

### Survival models

Cox and accelerated failure-time (AFT) models can share a table. An AFT model
reports time ratios, `exp()` of the log-time coefficients: a time ratio above
1 means longer survival, so it points the opposite way from a hazard ratio.

```{r survival}
d <- lung
d$sex <- factor(d$sex, 1:2, c("Male", "Female"))
d$dead <- as.integer(d$status == 2)
attr(d$sex, "label") <- "Sex"
attr(d$age, "label") <- "Age (years)"

cox <- coxph(Surv(time, dead) ~ sex + age, data = d, ties = "breslow")
weib <- survreg(Surv(time, dead) ~ sex + age, data = d, dist = "weibull")
regtab(cox, weib, models = c("Cox", "Weibull AFT"))
```

The Weibull shape parameters (`ln_p`, `p`, `1/p`) and the intercept appear
with `keepintercept = TRUE`, on their own scale. The lognormal and
log-logistic scale rows (`lnsigma`, `sigma`, `lngamma`, `gamma`) are kept
even without it, as in Stata.

For competing risks, the Fine-Gray subdistribution hazard ratio can come
from `coxph()` on `finegray()` data or from `cmprsk::crr()`. regtab
recognises a Fine-Gray `coxph()` fit from its structure (weighted, a
`Surv(start, stop, status)` response, data produced by `finegray()`) and
uses a sandwich variance clustered by subject; keep the subject id in the
`finegray()` call and pass it as `cluster`. `crr()` takes a covariate
matrix, so its rows are named by the matrix's column names.

```{r finegray, eval = has("cmprsk")}
m <- mgus2
m$etime <- with(m, ifelse(pstat == 0, futime, ptime))
m$event <- factor(with(m, ifelse(pstat == 0, 2 * death, 1)), 0:2,
                  c("censored", "progression", "death"))
m$id <- seq_len(nrow(m))
attr(m$age, "label") <- "Age at diagnosis"

fg <- finegray(Surv(etime, event) ~ age + sex + id, data = m,
               etype = "progression")
fg_cox <- coxph(Surv(fgstart, fgstop, fgstatus) ~ age + sex, data = fg,
                weights = fgwt, cluster = id)
fg_crr <- cmprsk::crr(m$etime, as.integer(m$event) - 1,
                      cov1 = cbind(age = m$age, male = as.integer(m$sex == "M")),
                      failcode = 1, cencode = 0)
regtab(fg_cox, fg_crr, models = c("coxph + finegray()", "cmprsk::crr()"))
```

The `crr()` column's `age` row joins the `coxph()` fit's age row because
their keys match (`age`); its `male` column has no factor structure, so it
gets its own row. `crr()` reproduces Stata's `stcrreg` exactly; the weighted
Cox fit approximates it.

### Ordinal and multinomial models

Ordered logistic models give odds ratios, with the cutpoints dropped
together with the intercept. `keepintercept = TRUE` shows them on the raw
scale, without p-values, and `cutlabels` names them. The outcome must be an
ordered factor, but a predictor must not be one (see
[What regtab refuses](#what-regtab-refuses)).

```{r ordinal, eval = has("MASS")}
h <- MASS::housing
h$Infl <- factor(h$Infl, ordered = FALSE)
h$Cont <- factor(h$Cont, ordered = FALSE)
attr(h$Infl, "label") <- "Perceived influence"
attr(h$Cont, "label") <- "Contact with other residents"
po <- MASS::polr(Sat ~ Infl + Cont, weights = Freq, data = h, Hess = TRUE)
regtab(po, keepintercept = TRUE, cutlabels = c("Low | Medium", "Medium | High"))
```

(`polr()` weights are frequency weights here, one row per cell of the
contingency table, which is what they are.)

A multinomial model has one equation per non-base outcome. Its rows read
`<outcome>: <term>`, and factor levels are written as Stata's keys
(`2.Infl`), as Stata's `mlogit` table shows them:

```{r multinom, eval = has("MASS", "nnet")}
hs <- h[rep(seq_len(nrow(h)), h$Freq), ]
hs$Sat <- factor(hs$Sat, ordered = FALSE)
mn <- nnet::multinom(Sat ~ Infl + Cont, data = hs, trace = FALSE)
regtab(mn)
```

### Mixed models

For `lmer()`, `glmer()` and `nlme::lme()` fits, the random-effects
parameters follow the fixed effects: variances, covariances and the
residual variance, with Wald intervals computed on the log standard
deviation. `relabel = TRUE` gives them readable names; `noreeffects = TRUE`
drops them. Fit linear mixed models with `REML = FALSE` (`lmer()`) or
`method = "ML"` (`lme()`) to match Stata's `mixed` (maximum likelihood).

```{r mixed, eval = has("lme4")}
ss <- lme4::sleepstudy
attr(ss$Days, "label") <- "Days of sleep deprivation"
mx <- lme4::lmer(Reaction ~ Days + (Days | Subject), data = ss, REML = FALSE)
regtab(mx, relabel = TRUE, stats = c("n", "groups", "icc"))
```

## Several models in one table

A simulated cohort is used from here on: myocardial infarction (MI) in
patients from 30 clinics, with smoking status as the exposure.

```{r cohort}
set.seed(20260926)
n <- 1200
clinic <- sample(1:30, n, replace = TRUE)
rural <- rbinom(30, 1, 0.4)
u <- rnorm(30, 0, 0.5)
cohort <- data.frame(
  clinic  = clinic,
  rural   = rural[clinic],
  smoking = factor(sample(c("Never", "Former", "Current"), n, TRUE,
                          c(0.5, 0.3, 0.2)),
                   levels = c("Never", "Former", "Current")),
  age     = round(rnorm(n, 60, 10)),
  bmi     = round(rnorm(n, 27, 4), 1)
)
lp <- with(cohort, -1.3 + u[clinic] + 0.4 * rural +
             0.4 * (smoking == "Former") + 0.9 * (smoking == "Current") +
             0.03 * (age - 60) + 0.05 * (bmi - 27))
cohort$mi <- rbinom(n, 1, plogis(lp))
labs <- c(smoking = "Smoking status", age = "Age (years)",
          bmi = "Body-mass index (kg/m2)", rural = "Rural clinic",
          mi = "Myocardial infarction")
for (v in names(labs)) attr(cohort[[v]], "label") <- labs[[v]]
```

Pass the models in order and name them with `models` (or pass a named list).
The rows are the union of all models' rows in order of first appearance,
with blank cells where a model lacks a term.

```{r multi}
crude <- glm(mi ~ smoking, family = binomial, data = cohort)
adjusted <- glm(mi ~ smoking + age + bmi, family = binomial, data = cohort)
full <- glm(mi ~ smoking + age + bmi + rural, family = binomial, data = cohort)
regtab(list(Crude = crude, Adjusted = adjusted, Full = full))
```

Factor levels are joined **by level**, not by position. A model with a
different reference level puts its Reference on that level's row, and a
model fitted on a subset that lacks a level leaves that row blank:

```{r join}
releveled <- cohort
releveled$smoking <- relevel(releveled$smoking, ref = "Current")
m_ref <- glm(mi ~ smoking + age, family = binomial, data = releveled)

no_former <- droplevels(subset(cohort, smoking != "Former"))
m_sub <- glm(mi ~ smoking + age, family = binomial, data = no_former)

regtab(adjusted, m_ref, m_sub,
       models = c("Never as reference", "Current as reference", "No former smokers"),
       keep = c("smoking", "age"))
```

Base R's `relevel()`, `subset()` and row indexing with `[` drop the
`"label"` attribute, and a row takes its label from the first model in which
it appears. Here that is the first model, fitted on the labelled data; put
the relabelled or subset model first, or table it alone, and its rows show
the column names until the labels are set again.

## Choosing and labelling rows

### keep and drop

`keep` and `drop` (not both) take Stata-style keys: a variable name selects
its label row, its levels and every interaction containing it; `2.smoking`
selects one level (its position, or its Stata value code for data read with
`tt_as_factor()`); `_cons` is the intercept. R coefficient names such as
`smokingCurrent` also work, and `labelmatch = TRUE` matches the displayed
labels as case-insensitive substrings instead.

```{r keep}
regtab(crude, adjusted, full, models = c("Crude", "Adjusted", "Full"),
       keep = "smoking")
```

A `keep` that matches nothing is an error. A `drop` that matches nothing is
silently ignored, as in Stata. The key of a term written as a function call
in the formula is the call itself, so with `factor(period)` in a formula it
is `drop = "factor(period)"`, not `"period"`. Converting the column to a
factor before fitting gives both a plain key and a readable header row.

### Labels

Row labels come from the `"label"` attribute of each variable, level rows
from the factor levels. Other text is set by arguments:

| Argument | Sets |
|---|---|
| `models` | the model headers |
| `coef` | the estimate header for every model (`"RR"`, `"aOR"`, ...) |
| `refcat`, `omitlabel`, `emptylabel` | the text of base-level, collinear and empty rows |
| `cutlabels` | ordinal cutpoint labels |
| `relabel` | readable random-effects labels |
| `title`, `footnote` | the title and footnote carried to every output |
| `xsymbol`, `vsref` | interaction labels (`interactions = "fvgen"`, the default) |

### Model statistics and extra rows

`stats` adds rows below the estimates, always in Stata's order: `n`
(Observations, or Subjects for survival models), `events` (failures of a
Cox, Fine-Gray or parametric survival model), `groups`, `mi_m` (the number
of imputations), `aic`, `qic`, `bic`, `ll`, `icc`, `r2` (R-squared for
linear models, the pseudo R-squared for the models Stata reports one for),
and for linear regression `r2_a` (adjusted R-squared), `rmse` (root mean
squared error) and `F` (the model F statistic, never shown for a logistic
or other generalized linear model), then `fmi` (the largest fraction of
missing information of a multiply imputed model). The R-only `vce` adds a
"Standard errors" row naming each model's variance ("model-based",
"robust", "robust, clustered by id", ...). `stat_fun` adds rows of
your own, computed from each fit. `addrow` adds
free-text rows, one value per model, such as a test for trend.

```{r stats}
# Test for trend: smoking entered as a score (0, 1, 2).
cohort$smoke_score <- as.integer(cohort$smoking) - 1L
p_trend <- c(
  coef(summary(glm(mi ~ smoke_score, binomial, cohort)))["smoke_score", 4],
  coef(summary(glm(mi ~ smoke_score + age + bmi, binomial, cohort)))["smoke_score", 4]
)
p_trend <- ifelse(p_trend < 0.001, "<0.001", sprintf("%.3f", p_trend))

regtab(crude, adjusted, models = c("Crude", "Adjusted"),
       stats = c("n", "aic", "bic", "r2"),
       addrow = list("P for trend" = p_trend))
```

Numbers given to `addrow` are printed in full decimal form, so format
p-values as text first, as above.

### Multiple imputation

A multiply imputed analysis is one model: a `mice` `mira`
(`with(imp, glm(...))`), or `tt_mi()` of a list of fits, one per completed
dataset. regtab pools it as Stata's `mi estimate` does (Rubin's rules, with
Barnard and Rubin's small-sample degrees of freedom for linear regression
and the large-sample ones for logistic and Cox models), takes the rows and
labels from the first imputation, and adds the Imputations and Largest FMI
rows on request. Here BMI is missing for a fifth of the cohort, and filled
in five times by a crude random draw (use `mice` in practice):

```{r mi}
miss <- runif(nrow(cohort)) < 0.2
imputed <- lapply(1:5, function(m) {
  d <- cohort
  d$bmi[miss] <- round(rnorm(sum(miss), mean(cohort$bmi), sd(cohort$bmi)), 1)
  glm(mi ~ smoking + age + bmi, family = binomial, data = d)
})
regtab(adjusted, tt_mi(imputed), models = c("Complete data", "Imputed BMI"),
       stats = c("n", "mi_m", "fmi"))
```

### Intercepts and display

The intercept is dropped by default when every model is on a ratio scale
(`nointercept = TRUE`) and kept otherwise; `keepintercept = TRUE` or
`nointercept = TRUE` overrides that. `digits` sets the decimals of estimates
and limits, `compact = TRUE` puts the estimate and interval in one column,
`nopvalue = TRUE` removes the p-values, `stars = TRUE` adds significance
stars, and `dimnonsig = TRUE` greys rows whose interval includes the null in
every model (in the workbook, Word and HTML output).

```{r display}
lin <- lm(bmi ~ smoking + age, data = cohort)
regtab(lin, digits = 3, compact = TRUE, stars = TRUE)
```

## Standard errors {#standard-errors}

`vce` chooses the variance behind the standard errors, intervals and
p-values:

| `vce` | Variance | Stata |
|---|---|---|
| `"stata"` (default) | what the Stata command reports by default: observed information for GLMs, OLS for `lm`, the model-based variance for GEE | default `vce(oim)`/`vce(ols)` |
| `"model"` | the fit's own `vcov()`, with R's `summary()` reference distribution | |
| `"robust"` | the sandwich with Stata's small-sample factor: HC0 times N/(N - 1) for GLMs (z), HC1 for `lm` (t) | `vce(robust)`, `[pweight]` |
| `"cluster"` | clustered on `cluster = ~id`, times G/(G - 1) (and (N - 1)/(N - k) for `lm`, with t on G - 1 degrees of freedom) | `vce(cluster id)` |

For most classes `"stata"` and `"model"` give the same numbers to display
precision; they differ, for example, for non-canonical links, where R's
`vcov()` uses the expected information. `vce` takes one value or one per
model, and `cluster` one variable or a list with one entry per model.
Giving `cluster` alone implies `vce = "cluster"`.

```{r vce}
regtab(full, full, full,
       models = c("Model-based", "Robust", "Clustered by clinic"),
       vce = c("stata", "robust", "cluster"), cluster = ~clinic,
       keep = c("smoking", "rural"), vce_note = TRUE)
```

Rural clinic is a clinic-level variable, so its clustered interval is much
wider; smoking varies within clinics and changes little. `vce_note = TRUE`
appends a sentence naming each model's non-default variance to the
footnote, which is written to the workbook, Word and HTML (the console
listing does not print it):

```{r vce_note}
tab <- regtab(full, full, full,
              models = c("Model-based", "Robust", "Clustered by clinic"),
              vce = c("stata", "robust", "cluster"), cluster = ~clinic,
              vce_note = TRUE)
tab$footnote
```

Not every class has a robust variance. `tt_vce_types()` lists the `vce`
values a fit accepts, and `tt_ci_methods()` the interval methods:

```{r vce-types}
tt_vce_types(full)
tt_vce_types(po)
tt_ci_methods(full)
tt_ci_methods(cox)
```

Confidence intervals are Wald intervals by default, consistent with the
p-values. `ci_method = "profile"` gives profile-likelihood intervals for
`lm`, `glm` and `glm.nb` fits with a model-based variance; it is refused with
a robust or clustered variance, whose intervals would then not match the
p-values.

`tt_vcov()` returns the matrix regtab uses, so other tools can share it. For
example, the modified Poisson risk ratio mentioned above, and a marginal
risk difference from the same robust variance with marginaleffects:

```{r modified-poisson}
mp <- glm(mi ~ smoking + age + bmi, family = poisson, data = cohort)
regtab(mp, vce = "robust", coef = "RR", keep = "smoking")
```

```{r marginaleffects, eval = has("marginaleffects")}
lg <- glm(mi ~ smoking + age + bmi, family = binomial, data = cohort)
marginaleffects::avg_comparisons(lg, variables = "smoking",
                                 vcov = tt_vcov(lg, "robust"))
```

Pass the matrix, not a string: marginaleffects' own `vcov = "stata"` is
HC1, which is not the variance Stata's `logit, vce(robust)` reports.
`effecttab()` (Stata's `effecttab`) lays such results out as a table; see
[Weighted analyses](weighted-analyses.html) for treatment effects and
margins.

## The numbers behind the table {#forest-plots}

`as_forest_data()` returns one row per displayed estimate and model, at full
precision on the display scale, plus the Reference rows (with missing
numbers). It is Stata's `regtab, eplotframe()`, and the input for a forest
plot:

```{r forest-data}
fd <- as_forest_data(regtab(crude, adjusted, models = c("Crude", "Adjusted"),
                            keep = "smoking"))
fd
```

```{r forest-plot, fig.width = 6, fig.height = 2.8, fig.alt = "Forest plot of odds ratios for former and current smoking, crude and adjusted."}
est <- fd[fd$rowtype == "effect", ]
y <- rev(seq_len(nrow(est)))
op <- par(mar = c(4, 12, 1, 1))
plot(est$estimate, y, log = "x", xlim = range(est$ll, est$ul, 1), pch = 15,
     yaxt = "n", ylab = "", xlab = "Odds ratio (95% CI)")
segments(est$ll, y, est$ul, y)
abline(v = 1, lty = 2)
axis(2, at = y, labels = paste(trimws(est$label), est$model_label, sep = ", "),
     las = 1)
par(op)
```

`as.data.frame(tab)` gives the table itself as text cells, as it is written
to CSV.

## A cohort-study Table 2: rates beside hazard ratios {#table-2}

The usual second table of a cohort study shows, for each exposure category,
the number of events, the person-time and the incidence rate, next to the
hazard ratio from a Cox model. `stratetab()` builds the rate part and
`regtab()` the models; `hrcomptab()` (Stata's `hrcomptab`) puts the selected
model rows beside the rates, each on the category with the same label,
and marks the category left over as the reference.

```{r table2-data}
lung2 <- lung
lung2$dead <- as.integer(lung2$status == 2)
lung2$years <- lung2$time / 365.25
lung2$sex <- factor(lung2$sex, 1:2, c("Male", "Female"))
lung2$ecog <- factor(pmin(lung2$ph.ecog, 2), 0:2, c("0", "1", "2 or more"))
lung2 <- lung2[!is.na(lung2$ecog), ]
attr(lung2$sex, "label") <- "Sex"
attr(lung2$ecog, "label") <- "ECOG performance status"
attr(lung2$age, "label") <- "Age (years)"

# Events, person-years and deaths per 100 person-years, by sex and by ECOG
# status: one block per exposure. outcomeids names the event variable, which
# hrcomptab() matches to the outcome of each coxph() model.
rates <- stratetab(list(tt_rates(lung2, "years", "dead", by = "sex"),
                        tt_rates(lung2, "years", "dead", by = "ecog")),
                   outcomes = 1, outlabels = "Death", outcomeids = "dead",
                   explabels = c("Sex", "ECOG performance status"),
                   ratescale = 100, unitlabel = "100")
```

The adjusted model holds both exposures. Its non-reference rows are
selected by label (`rownames`) or by row number (`rows`); they are taken
block by block in the rate table's order (one row for the two sexes, two for
the three ECOG groups) and placed by label:

```{r table2-adjusted}
adj <- regtab(coxph(Surv(years, dead) ~ sex + ecog + age, data = lung2,
                    ties = "breslow"), coef = "HR")
adj

hrcomptab(rates, adj, rownames = c("female", "1", "2 or more"),
          title = "Table 2. Deaths by sex and ECOG status",
          footnote = "aHR: hazard ratio adjusted for sex, ECOG status and age.")
```

`rownames` are case-insensitive patterns matched as *parts* of the row
labels, with `*` and `?` as wildcards: `"female"` selects Female only, while
`"male"` would select both Male and Female (as `"treated"` would match
Untreated). When one label contains another, match whole labels with
`rownames_exact = TRUE` (R only; `"male"` then selects Male alone), or
select by row number with `rows`, as below.

Crude hazard ratios come from one model per exposure, so there are two model
tables, one selection each:

```{r table2-crude}
crude_sex <- regtab(coxph(Surv(years, dead) ~ sex, data = lung2, ties = "breslow"),
                    coef = "HR")
crude_ecog <- regtab(coxph(Surv(years, dead) ~ ecog, data = lung2, ties = "breslow"),
                     coef = "HR")
hrcomptab(rates, list(crude_sex, crude_ecog), rows = list(3, 3:4), effect = "HR")
```

The models are matched to the rate table's outcomes by identity, never by
position: the rate table's `outcomeids` against each model's outcome (for a
`coxph()` fit, the event variable). With several outcomes, each model table
holds one model per outcome (`regtab(list(fit_death, fit_relapse))`), and
`outcomemap` names the model for each outcome when the identities differ
(model labels will do). Anything that cannot be placed exactly -- a heading
or reference row, a level no category is labelled with, a reference category
that is not the model's base level, models at another confidence level or on
another scale -- is an error rather than a guess.

`comptab()` without a rate table stacks selected rows of several model
tables, here the crude and adjusted estimates for sex, with section headings
and the estimate and interval in one column:

```{r table2-vertical}
comptab(list(crude_sex, adj), rownames = list("female", "female"),
        section = c("Crude", "Adjusted for ECOG status and age"), compact = TRUE)
```

Both write the workbook, CSV and Markdown files like the other commands
(`xlsx =`, `csv =`, `markdown =`, after which the table is returned
invisibly: assign it to print it), convert with `as_flextable()` and
`tt_as_gt()`, and give their estimates to `as_forest_data()`, with the
section and reference rows of the composite.

## Models regtab does not support: data-frame input {#data-frame-input}

regtab accepts only the classes listed above; any other model, including a
subclass of a supported one, is refused with an error that names it. Such a
model can still be tabled: pass a data frame shaped like
`broom.helpers::tidy_plus_plus()` output. `survey::svycoxph()`, a Cox model
for survey data, is one example:

```{r svycoxph-refused, error = TRUE, eval = has("survey")}
set.seed(1)
d$w <- runif(nrow(d), 0.5, 2)
design <- survey::svydesign(ids = ~1, weights = ~w, data = d)
svy_cox <- survey::svycoxph(Surv(time, dead) ~ sex + age, design = design)
regtab(svy_cox)
```

```{r svycoxph-tidy, eval = has("survey"), results = "hide"}
svy_df <- broom.helpers::tidy_plus_plus(svy_cox, exponentiate = TRUE)
```

```{r svycoxph-df, eval = has("survey")}
regtab(cox, svy_df, models = c("coxph, unweighted", "svycoxph"))
```

The data frame joined the fitted model's rows because `tidy_plus_plus()`
records the variable, its levels and the reference level, and regtab keys
them the same way (`2.sex`). `tidy_plus_plus()` also records that it
exponentiated the estimates, which sets the header (`HR`) and the null
value.

regtab takes a data frame's numbers as they are and never transforms them.
Two things it cannot know, and will not guess:

- **The effect scale.** Whether `estimate` holds ratios or coefficients
  sets the header, the null value (1 or 0) for derived p-values and for
  `dimnonsig`, and whether the intercept is dropped. It is read from
  `attr(x, "effect_scale")` (`"OR"`, `"HR"`, `"RR"`, `"Coef."`, ...), else
  from `attr(x, "stata_cmd")`, else from broom.helpers' record of
  exponentiating. Output of `broom::tidy(exponentiate = TRUE)` with no
  declared scale is refused: its `statistic` column shows that the
  estimates were exponentiated, and taking them as coefficients would test
  them against 0.
- **The scale of `std.error`.** When a ratio row has no interval (or the
  data frame has no p-values), regtab derives them from `std.error`. For
  broom, `std.error` is the standard error of log(estimate); for the
  parameters package it is the standard error of the ratio itself (delta
  method). The two give different intervals and p-values, so on a ratio
  scale regtab requires `attr(x, "se_scale")`: `"link"` or `"estimate"`.

```{r se-scale, error = TRUE}
b <- coef(cox)
se <- sqrt(diag(vcov(cox)))
hr <- data.frame(term = names(b), estimate = exp(b), std.error = se)
attr(hr, "effect_scale") <- "HR"
regtab(hr)

attr(hr, "se_scale") <- "link"
regtab(hr)
```

Declaring the wrong scale gives wrong intervals (here `"estimate"` would
put the interval at the hazard ratio plus or minus 1.96 times the standard
error of its logarithm), which is why regtab asks rather than guesses.
Supplied `conf.low`/`conf.high` and `p.value` columns are shown as they
are; note that `tidy_plus_plus()` computes a `glm`'s intervals by profile
likelihood (from `confint()`), so a data frame from a `glm` can differ from
regtab's Wald interval for the same fit in the last digit. `?regtab`,
section "Data-frame input", lists every column and attribute.

## What regtab refuses {#what-regtab-refuses}

regtab refuses a model or an option rather than print a table whose rows
could be misread. Each error names the problem and, where there is one, the
alternative.

- **Unsupported classes and subclasses** (`mgcv::gam`,
  `survey::svycoxph`, penalised or frailty `coxph`, mixed-effects Cox
  models from `coxme`): their variance has not been checked against a
  Stata command. Pass a data frame, as above.
- **Ordered factors and non-treatment contrasts as predictors**: a
  Reference row claims a contrast against the base level, which polynomial,
  sum or Helmert contrasts do not estimate. Convert the predictor with
  `factor(x, ordered = FALSE)`.
- **A variance a class does not have**: `vce = "robust"` or `"cluster"` for
  `polr`, `clm`, `multinom`, `zeroinfl`, `hurdle` and mixed models, and any
  `vce` outside the four names (there is no `"HC3"`).
- **Profile intervals beside a robust variance**, or for classes other
  than `lm`, `glm` and `glm.nb`.
- **A pooled `mice` result** (`mipo`): its degrees of freedom are not
  Stata's `mi estimate`'s. Pass the `mira` (or `tt_mi()` of the fits), which
  regtab pools itself (see "Multiple imputation" above); `fit$analyses` as
  a list shows the imputations side by side. Imputed fits of classes other
  than `lm`, `glm`, `coxph` and `crr` are refused.
- **Fits whose data changed**: a fit that keeps no copy of its data
  (`coxph`, `survreg`, `multinom`, `model = FALSE` fits) is refused when the
  data it was fitted on were filtered or edited afterwards, since its
  labels, reference rows and N would come from other data than its
  estimates. Refit, or fit with `model = TRUE`.

```{r refusals, error = TRUE}
mt <- mtcars
mt$cyl <- factor(mt$cyl, ordered = TRUE)
regtab(lm(mpg ~ cyl, data = mt))

regtab(po, vce = "robust")

regtab(full, vce = "robust", ci_method = "profile")
```

Weighted fits are not refused, but regtab does not guess what kind of
weight they carry; see
[Weighted analyses: IPTW and marginal structural models](weighted-analyses.html).
For the Stata command each R model stands for, and every deliberate
difference from Stata's output, see [Coming from Stata](coming-from-stata.html).
