---
title: "Get started with tabtools"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with tabtools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
# The examples use survival (Suggests): without it the code is shown, not
# run.
has_survival <- requireNamespace("survival", quietly = TRUE)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = has_survival)
has_word <- has_survival && requireNamespace("flextable", quietly = TRUE) &&
  requireNamespace("officer", quietly = TRUE)
```

tabtools makes the two tables almost every clinical or epidemiological paper
needs: a table of baseline characteristics ("Table 1") and a table of
regression estimates. `table1_tc()` (alias `desctab()`) builds the first,
`regtab()` the second. Both return a `tt_table` object that prints in the
console and can be written to Excel, CSV and Markdown, or converted for Word
and HTML, with the same rows in every output.

The same package also makes incidence-rate tables, a cohort study's "Table
2" of rates beside hazard ratios, treatment-effect tables and
weight diagnostics; [Where next](#where-next) points to them. tabtools is the
R implementation of the Stata `tabtools` commands, but you do not need to
know Stata to use it; if you do, see [Coming from Stata](coming-from-stata.html).

## Installation

```{r install, eval = FALSE}
# install.packages("remotes")
remotes::install_github("tpcopeland/R-tabtools")
```

## Data

The examples use `survival::lung`, the NCCTG advanced lung cancer cohort.
Two habits make the tables read well:

- **Variable labels** are taken from the `"label"` attribute of each column,
  the attribute `haven::read_dta()` and the labelled packages set. Without
  one, the column name is shown.
- **Categorical variables** should be factors; the level order is the row
  order, and in regression tables the first level is the reference.

```{r data}
library(tabtools)
library(survival)

d <- lung[!is.na(lung$ph.ecog), ]
d$sex <- factor(d$sex, 1:2, c("Male", "Female"))
d$ecog <- factor(pmin(d$ph.ecog, 2), 0:2,
                 c("Asymptomatic", "Symptomatic, ambulatory",
                   "In bed part of the day"))
d$dead <- as.integer(d$status == 2)
d$karno_low <- as.integer(d$ph.karno < 80)

labs <- c(age = "Age (years)", sex = "Sex", ecog = "ECOG performance status",
          karno_low = "Karnofsky score below 80",
          meal.cal = "Calories consumed at meals",
          wt.loss = "Weight loss, last 6 months (lb)")
for (v in names(labs)) attr(d[[v]], "label") <- labs[[v]]
```

## Table 1

`vars` names the rows and, optionally, how each is summarised: `contn` mean
and SD (with an optional display format), `conts` median and quartiles, `cat`
counts and column percentages for each level, `bin` one row for the count of
1s. `by` splits the columns, `total` adds an overall column, and `smd`
adds standardized mean differences between the first two groups.

```{r table1}
t1 <- table1_tc(d, by = "sex",
                vars = c(age = "contn %5.1f", ecog = "cat", karno_low = "bin",
                         meal.cal = "conts", wt.loss = "conts"),
                total = "before", smd = TRUE)
t1
```

P-values come from R's conventional tests: Welch's t test for `contn`,
Wilcoxon rank-sum for `conts`, and Pearson's chi-squared test without
continuity correction for `cat` and `bin` (`cate` and `bine` give Fisher's
exact test). `test = TRUE` adds a column naming the test, `nopvalue = TRUE`
removes the p-values, and `test_args` passes arguments to the test
functions. A methods sentence describing the table is stored with it:

```{r table1-methods}
t1$stored$methods
```

Give `vars` as a plain character vector and each variable is typed
automatically: factors and few-valued numbers become `cat`, 0/1 variables
`bin`, and other numbers `contn` or `conts` depending on a normality test. Naming the types, as above, keeps the table
stable when the data change. `?table1_tc` lists every option: missing-value
rows, weights, small-cell suppression, number formats and cell delimiters.

## A regression table

`regtab()` takes one or more fitted models and puts them side by side. The
effect measure is chosen from the model: hazard ratios for a Cox model, odds
ratios for a logistic model, incidence rate ratios for a Poisson model, and
so on, exponentiated where the model is on a ratio scale.

```{r regtab}
# ties = "breslow" reproduces Stata's numbers. R's default (Efron) is valid,
# and a better approximation with tied times; regtab then notes once that
# Stata's stcox (Breslow) gives slightly different estimates.
crude <- coxph(Surv(time, dead) ~ sex, data = d, ties = "breslow")
adjusted <- coxph(Surv(time, dead) ~ sex + age + ecog, data = d,
                  ties = "breslow")

t2 <- regtab(crude, adjusted, models = c("Crude", "Adjusted"),
             stats = c("n", "aic"))
t2
```

Factor variables get a label row, indented level rows, and a Reference row
for the base level. Rows are matched across models by variable and level, so
a covariate that appears in only one model has blank cells in the other.
Model statistics (`stats`) go below the estimates. As for Table 1, a methods
sentence is stored:

```{r regtab-methods}
t2$stored$methods
```

[Regression tables with regtab()](regression-tables.html) covers the
supported model classes, the row options, and robust and clustered standard
errors.

## Writing the tables

The `xlsx`, `csv` and `markdown` arguments write a table as it is built; the
`tt_write_*()` functions write one later. Several tables can go to one
workbook as separate sheets. A command that writes a file returns its table
invisibly, so nothing prints: assign it (`t2 <- regtab(..., xlsx = xlsx)`)
and print it to see it.

```{r write}
out <- tempfile("tables")
dir.create(out)
xlsx <- file.path(out, "tables.xlsx")

tt_write_xlsx(t1, xlsx, sheet = "Table 1")
tt_write_xlsx(t2, xlsx, sheet = "Table 2")
tt_write_csv(t2, file.path(out, "table2.csv"))
tt_write_markdown(t2, file.path(out, "table2.md"))

# The same at build time (nothing prints):
regtab(crude, adjusted, models = c("Crude", "Adjusted"),
       xlsx = xlsx, sheet = "Table 2b")

list.files(out)
openxlsx2::wb_get_sheet_names(openxlsx2::wb_load(xlsx))
```

The workbook carries the house style: bold headers, merged model headers,
italic Reference cells spanning the estimate, interval and p-value columns,
and borders set by `borderstyle`. The CSV keeps the label indents as leading
spaces; the Markdown file writes them as `&nbsp;` so they survive rendering:

```{r markdown-file}
cat(readLines(file.path(out, "table2.md")), sep = "\n")
```

`as.data.frame()` returns the table as a data frame of text cells, and
`as_forest_data()` the numbers behind a regression table (see
[Regression tables with regtab()](regression-tables.html#forest-plots)).

## Word and HTML

`flextable::as_flextable()` converts a table to a flextable with the same
house style as the workbook. In an R Markdown or Quarto document it renders
directly:

```{r flextable, eval = has_word}
flextable::as_flextable(t2)
```

For a Word file, add the flextables to an officer document:

```{r docx, eval = has_word}
doc <- officer::read_docx()
doc <- flextable::body_add_flextable(doc, flextable::as_flextable(t1))
doc <- officer::body_add_break(doc)
doc <- flextable::body_add_flextable(doc, flextable::as_flextable(t2))
docx <- file.path(out, "tables.docx")
print(doc, target = docx)
file.exists(docx)
```

`flextable::save_as_docx(flextable::as_flextable(t2), path = ...)` writes a
single table. `title` and `footnote` (arguments of both commands) become the
caption and footer line. For gt output, used in HTML and Quarto, see
`tt_as_gt()`.

## Settings

`tabtools_options()` sets defaults for the font, font size, border style,
digits and colours, for the session or, with `persist = TRUE`, for future
sessions:

```{r options, eval = FALSE}
tabtools_options(font = "Calibri", fontsize = 11, borderstyle = "academic")
```

## Where next {#where-next}

- [Regression tables with regtab()](regression-tables.html): model classes
  and their effect measures, multi-model tables, `keep`/`drop`, standard
  errors, tables from models regtab does not support, and
  [a Table 2 of incidence rates beside hazard ratios](regression-tables.html#table-2)
  (`tt_rates()`, `stratetab()`, `hrcomptab()`).
- [Weighted analyses: IPTW and marginal structural models](weighted-analyses.html):
  balance tables before and after weighting, weight diagnostics
  (`wttab()`), robust and cluster-robust standard errors, marginal
  structural models, and treatment effects and margins (`effecttab()`).
- [Compared with gtsummary](compared-with-gtsummary.html): the same two
  tables made with gtsummary and with tabtools.
- [Coming from Stata](coming-from-stata.html): the Stata to R translation,
  option by option, and the deliberate differences.
