---
title: "Weighted analyses: IPTW and marginal structural models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Weighted analyses: IPTW and marginal structural models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
# Every example uses survival (Suggests): without it the code is shown,
# not run.
has <- function(...) all(vapply(c("survival", ...), requireNamespace, TRUE, quietly = TRUE))
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = has())
```

A paper that uses inverse probability of treatment weighting (IPTW), or a
marginal structural model (MSM) fitted with time-varying weights, usually
reports four things:

1. covariate balance before and after weighting, as standardized mean
   differences (SMDs), often with the effective sample size;
2. the distribution of the weights, and how they were truncated, if they
   were;
3. the weighted effect estimate (a marginal odds, hazard or risk ratio, or
   a risk difference) with a confidence interval that accounts for the
   weighting, often beside crude and conventionally adjusted estimates;
4. sensitivity analyses, such as truncated weights.

tabtools makes a table for each: `table1_tc()` for (1); `wttab()` for (2),
the distribution of the weights overall, by treatment group and by
follow-up period, with the effective sample size and the effect of
truncation; `regtab()` for (3) and (4); and `effecttab()` for treatment
effects, standardized risks and risk differences computed by
marginaleffects.

The one thing to get right with weighted models in R is the variance. R's
`glm(weights = w)` does not know whether `w` is a frequency weight, an
analytic weight or an inverse-probability weight. Stata does: `[pweight]`
always reports a robust variance. regtab therefore never guesses; the `vce`
argument says which variance a model gets.

```{r setup}
library(tabtools)
library(survival)
```

## Part 1: IPTW for a point treatment

### Data and weights

A simulated cohort of 2,000 patients starting one of two drugs. Older,
frailer patients and those with chronic kidney disease (CKD) are more
likely to get the new drug and also more likely to die. Follow-up is three
years.

```{r cohort}
set.seed(20260926)
n <- 2000
d <- data.frame(
  age     = round(rnorm(n, 62, 10)),
  female  = rbinom(n, 1, 0.55),
  ckd     = rbinom(n, 1, 0.25),
  frailty = factor(sample(c("Low", "Medium", "High"), n, TRUE,
                          c(0.5, 0.3, 0.2)),
                   levels = c("Low", "Medium", "High"))
)
lp <- with(d, -0.5 + 0.04 * (age - 62) - 0.3 * female + 0.8 * ckd +
             0.5 * (frailty == "Medium") + 1.0 * (frailty == "High"))
d$treat <- factor(rbinom(n, 1, plogis(lp)), 0:1, c("Standard", "New drug"))
hazard <- with(d, 0.15 * exp(0.04 * (age - 62) + 0.6 * ckd +
                               0.4 * (frailty == "Medium") +
                               0.8 * (frailty == "High") -
                               0.35 * (treat == "New drug")))
t_event <- rexp(n, hazard)
d$time <- pmin(t_event, 3)
d$died <- as.integer(t_event <= 3)
d$died_1y <- as.integer(t_event <= 1)

labs <- c(age = "Age (years)", female = "Female sex",
          ckd = "Chronic kidney disease", frailty = "Frailty",
          treat = "Treatment", died_1y = "Death within 1 year")
for (v in names(labs)) attr(d[[v]], "label") <- labs[[v]]
```

Stabilized ATE weights from a logistic propensity score model: the marginal
probability of the treatment received over its conditional probability.

```{r weights}
ps <- glm(treat ~ age + female + ckd + frailty, family = binomial,
          data = d)$fitted.values
p_new <- mean(d$treat == "New drug")
d$sw <- ifelse(d$treat == "New drug", p_new / ps, (1 - p_new) / (1 - ps))
```

The WeightIt package gives the same weights and, in addition, an outcome
model whose variance accounts for estimating them (used below):

```{r weightit, eval = has("WeightIt"), message = FALSE}
W <- WeightIt::weightit(treat ~ age + female + ckd + frailty, data = d,
                        method = "glm", estimand = "ATE", stabilize = TRUE)
all.equal(unname(W$weights), d$sw)
```

### The weights

`wttab()` gives the usual summaries: the distribution overall and by
treatment group (N, mean, SD, minimum, percentiles, maximum) and Kish's
effective sample size, ESS = (sum w)^2 / sum w^2, also as a percentage of
N. The percentiles are Stata's (`_pctile`), and the table writes to a
workbook, Word or HTML like any other tabtools table.

```{r weight-summary}
wttab(d, "sw", by = "treat", digits = 3, title = "Stabilized weights")
```

A WeightIt object can be passed directly; its treatment gives the groups:

```{r weight-summary-weightit, eval = has("WeightIt")}
wttab(W, digits = 3)
```

Stabilized weights should average about 1 in each group. The effective
sample size within a group does not depend on stabilization: stabilizing
multiplies every weight in a group by the same constant, which cancels in
the ratio.

### Balance: Table 1 before and after weighting

`wt` names the weight variable. With `wtcompare = TRUE` the unweighted and
weighted columns sit side by side; `smd = TRUE` adds the SMD of the weighted
comparison; `wtn = TRUE` shows the weighted cells as effective counts,
(sum of weights in the cell / sum of weights in the group) times the group's
N, instead of percentages alone.

```{r balance}
vars <- c(age = "contn %5.1f", female = "bin", ckd = "bin", frailty = "cat")
bal <- table1_tc(d, by = "treat", vars = vars, wt = "sw", smd = TRUE,
                 wtcompare = TRUE, wtn = TRUE)
bal
```

Under `wt`, the N= row counts records and an "Effective sample size" row
follows it. P-values and tests are not computed under `wt`, as in Stata;
balance is judged from the SMDs. SMDs
above `smdthreshold` (default 0.1) are highlighted in the workbook, Word
and HTML output.

The SMD column of a `wtcompare` table is the weighted one. For the SMDs
before weighting, build the unweighted table as well; the SMDs of both are
stored in `$stored$table`, which also gives the numbers for a balance
("Love") plot:

```{r smd-before-after}
unw <- table1_tc(d, by = "treat", vars = vars, smd = TRUE, nopvalue = TRUE)
smd <- data.frame(unweighted = unw$stored$table[, "smd"],
                  weighted = bal$stored$table[, "smd"],
                  row.names = labs[names(vars)])
round(smd, 3)
```

```{r love-plot, fig.width = 6, fig.height = 2.6, fig.alt = "Standardized mean differences for four covariates before and after weighting; all fall below 0.1 after weighting."}
op <- par(mar = c(4, 11, 1, 1))
dotchart(smd$unweighted, labels = rownames(smd), xlim = c(0, 0.6), pch = 1,
         xlab = "Absolute standardized mean difference")
points(smd$weighted, seq_len(nrow(smd)), pch = 19)
abline(v = 0.1, lty = 2)
legend("topright", c("Unweighted", "Weighted"), pch = c(1, 19), bty = "n")
par(op)
```

### The outcome model and its variance

The weighted outcome model regresses the outcome on treatment alone (or
with the covariates, for a doubly robust estimate). With `glm()` the
weights enter as `weights = sw`; binomial fits then warn about non-integer
successes, which is harmless here.

```{r iptw-fit}
iptw <- suppressWarnings(
  glm(died_1y ~ treat, family = binomial, data = d, weights = sw)
)
```

Under the default `vce = "stata"`, a weighted `glm()` gets Stata's
analytic-weight (`[aweight]`) variance, which is not the IPTW variance.
regtab says so when the weights look like inverse-probability weights
(non-integer and not constant), but does not switch by itself:

```{r iptw-default}
regtab(iptw)
```

`vce = "robust"` is the variance Stata reports for `logit ... [pweight = sw]`:
the sandwich estimator times N/(N - 1). Here it is shown beside the crude and
conventionally adjusted models, which keep their model-based variance:

```{r iptw-table}
crude <- glm(died_1y ~ treat, family = binomial, data = d)
adjusted <- glm(died_1y ~ treat + age + female + ckd + frailty,
                family = binomial, data = d)
t_or <- regtab(crude, adjusted, iptw,
               models = c("Crude", "Adjusted", "IPTW"),
               vce = c("stata", "stata", "robust"),
               keep = "treat", stats = "n", vce_note = TRUE)
t_or
t_or$footnote
```

`vce_note = TRUE` names each model's non-default variance in the footnote,
which goes to the workbook, Word and HTML output. The robust sandwich treats
the weights as known; for the ATE this is usually conservative, because
it ignores the estimation of the propensity score. `WeightIt::glm_weightit()` fits the same
weighted model with an M-estimation variance that accounts for it, and
regtab uses that variance as it is:

```{r weightit-table, eval = has("WeightIt")}
iptw_m <- WeightIt::glm_weightit(died_1y ~ treat, data = d, weightit = W,
                                 family = binomial)
regtab(iptw, iptw_m, models = c("glm()", "glm_weightit()"),
       vce = c("robust", "stata"), keep = "treat", vce_note = TRUE)$footnote
regtab(iptw, iptw_m, models = c("glm()", "glm_weightit()"),
       vce = c("robust", "stata"), keep = "treat")
```

For `glm_weightit()` fits `vce = "stata"` and `"model"` both give the
M-estimation variance; `"robust"` and `"cluster"` are not offered
(`tt_vce_types(iptw_m)`).

### Risk ratios, risk differences and hazard ratios

The odds ratio is not collapsible, so the marginal odds ratio from IPTW
differs from the conditional one even without confounding. Two other
measures are common:

- a **risk ratio** from a weighted Poisson model with the robust variance
  (the log-binomial model is shown on the log scale by regtab, as in
  Stata; see [Regression tables](regression-tables.html)), with the header
  relabelled `RR`;
- a **risk difference** from a weighted linear model: an `lm()` with
  `vce = "robust"` (HC1 and t, as Stata's `regress [pweight]`), or a
  Gaussian `glm_weightit()`.

```{r rr-rd}
rr <- suppressWarnings(glm(died_1y ~ treat, family = poisson, data = d,
                           weights = sw))
rd <- lm(died_1y ~ treat, data = d, weights = sw)
regtab(rr, vce = "robust", coef = "RR", keep = "treat")
regtab(rd, vce = "robust", coef = "RD", digits = 3, keep = "treat")
```

For a weighted Cox model, survival computes a robust variance by itself
whenever the weights are not integers, and regtab applies Stata's G/(G - 1)
factor to it, as Stata's `stcox` after `stset [pweight]` reports it. No `vce`
argument is needed:

```{r cox}
# ties = "breslow": Stata's stcox refuses Efron ties with weights, so
# regtab notes that a weighted Efron fit has no Stata analogue.
cox_crude <- coxph(Surv(time, died) ~ treat, data = d, ties = "breslow")
cox_adj <- coxph(Surv(time, died) ~ treat + age + female + ckd + frailty,
                 data = d, ties = "breslow")
cox_iptw <- coxph(Surv(time, died) ~ treat, data = d, weights = sw,
                  ties = "breslow")
t_hr <- regtab(cox_crude, cox_adj, cox_iptw,
               models = c("Crude", "Adjusted", "IPTW"),
               keep = "treat", vce_note = TRUE)
t_hr
t_hr$footnote
```

### teffects and margins in R

Stata reports treatment effects with `teffects` and predictive margins or
marginal effects with `margins`, and tabtools formats either collection with
`effecttab`. R's `effecttab()` formats the same results from the
[marginaleffects](https://marginaleffects.com) package (or from a data frame
or a matrix). Like Stata's command, it fits nothing itself: it takes the
estimates, intervals and p-values and lays them out as the other tables.

#### Treatment effects: the ATE with its potential-outcome mean

`teffects ipw (y) (treat x...), ate` reports the average treatment effect
(ATE) and the potential-outcome mean of the control group. In R the IPW
estimate is a weighted outcome model, and marginaleffects gives both rows:
`avg_comparisons()` the contrast, `avg_predictions()` with the treatment set
to the control level the potential-outcome mean. Passed together in one
list, the two form one model, which `effecttab()` shows as Stata does:

```{r effecttab-ate, eval = has("marginaleffects", "WeightIt")}
fit_rd <- WeightIt::glm_weightit(died_1y ~ treat, data = d, weightit = W)
ate <- marginaleffects::avg_comparisons(fit_rd, variables = "treat")
po <- marginaleffects::avg_predictions(fit_rd,
                                       variables = list(treat = "Standard"))
effecttab(list(ate, po), clean = TRUE, effect = "RD", digits = 3,
          title = "Risk difference of death within 1 year (IPTW)")
```

Without `clean` the rows show Stata's raw keys (`r2vs1.treat`, `1.treat`);
`tlabels = c("Standard" = "Old drug")` relabels a level. The Gaussian
`glm_weightit()` fit carries the M-estimation variance that accounts for
estimating the weights, which is the variance `teffects ipw` reports, and
`effecttab()` reads the estimator (IPW) from it for the methods sentence.
On the Stata demo cohort and on Stata's `cattaneo2` data the table agrees
with Stata's at every displayed digit. `avg_predictions(fit,
variables = "treat")` gives both potential-outcome means, as
`teffects ..., pomeans` does.

#### Risk ratios

The recommended route to a risk ratio is the log ratio, exponentiated
(`comparison = "lnratioavg", transform = exp`): the p-value is the Wald test
of the log ratio against 0, that is of the ratio against 1, and the interval
is the exponentiated log-scale interval, so it stays positive and agrees
with the test. This is how ratios are reported after Stata's `teffects`
(`nlcom` or `margins` on the log scale, then exponentiated).

```{r effecttab-rr, eval = has("marginaleffects", "WeightIt")}
rr <- marginaleffects::avg_comparisons(fit_rd, variables = "treat",
                                       comparison = "lnratioavg",
                                       transform = exp)
effecttab(list(rr, po), clean = TRUE, effect = "RR", digits = 3)
```

A plain ratio (`comparison = "ratio"`) has a delta-method standard error on
the ratio scale, and marginaleffects tests it against 0. `effecttab()`
recomputes that p-value against 1, keeps the interval, and says so in the
footnote; the log-ratio route above needs no such repair.

#### Several estimators side by side

Several estimators go side by side as models, one list per model:

```{r effecttab-models, eval = has("marginaleffects", "WeightIt")}
fit_ra <- lm(died_1y ~ treat * (age + female + ckd + frailty), data = d)
ra <- list(marginaleffects::avg_comparisons(fit_ra, variables = "treat",
                                            vcov = "HC0"),
           marginaleffects::avg_predictions(fit_ra,
                                            variables = list(treat = "Standard"),
                                            vcov = "HC0"))
effecttab(list(IPW = list(ate, po), RA = ra), clean = TRUE, effect = "RD",
          digits = 3)
```

The point estimates of regression adjustment agree with `teffects ra`. Its
standard errors do too once the variance is robust: `teffects` reports a
sandwich variance, and `lm()`'s default model-based variance is narrower (on
Stata's `cattaneo2` example, 1.8% smaller for the ATE). With
`vcov = "HC0"` the gap is about 0.1%, the part `teffects` adds for the
sampling variation of the covariates, which the delta method treats as
fixed. `effecttab()` says so once when an `lm()`/`glm()` result arrives with
the default variance. AIPW and matching have no marginaleffects counterpart
with Stata's variance; pass their estimates as a data frame.

#### Margins: predictive margins and marginal effects

`margins treat` after a logistic model gives the predictive margins,
`margins, dydx(...)` the average marginal effects. In a margins table a
factor is headed by its variable label with its levels beneath, and a
comparison's base level is the Reference row, as `dydx(i.treat)` shows it:

```{r effecttab-margins, eval = has("marginaleffects")}
fit_lr <- glm(died_1y ~ treat + age + female + ckd + frailty,
              family = binomial, data = d)
effecttab(
  marginaleffects::avg_predictions(fit_lr, variables = "treat"),
  marginaleffects::avg_comparisons(fit_lr, variables = c("treat", "frailty")),
  models = c("Predicted risk", "Risk difference"), effect = "Risk",
  digits = 3
)
```

A 0/1 variable that is not a factor (`female` here) is a discrete change in
marginaleffects, also under `avg_slopes()`: that is Stata's `dydx(i.female)`.
Stata's `dydx(female)` without `i.` is the derivative, which can be far from
it (about 9% for a logit coefficient of 1.5 at a baseline risk near 30%); use
`dydx(i.female)` in Stata for the same number.

After a weighted model, pass the variance regtab uses and the estimation
weights, as Stata's `margins` averages with them (`margins` after
`logit [pweight]`; the two agree at every displayed digit):

```{r effecttab-weighted, eval = has("marginaleffects")}
effecttab(marginaleffects::avg_comparisons(iptw, variables = "treat",
                                           vcov = tt_vcov(iptw, "robust"),
                                           wts = d$sw),
          effect = "RD", digits = 3)
```

Do not pass `vcov = "stata"` to marginaleffects: there it means HC1, which
is not Stata's `logit [pweight]` variance.

#### The Stata to R mapping

| Stata (collected) | R input to `effecttab()` | Agreement |
|---|---|---|
| `teffects ipw (y) (t x), ate` | `glm_weightit(y ~ t, weightit = weightit(t ~ x))`; `list(avg_comparisons(fit, variables = "t"), avg_predictions(fit, variables = list(t = <control>)))` | estimates and M-estimation SEs, every displayed digit |
| `teffects ipw ..., atet` | the same with `weightit(..., estimand = "ATT")` | estimates and SEs (to about 1e-7) |
| `teffects ra ..., atet` | `lm(y ~ t * (x))` with `newdata = subset(d, t == 1)` | estimates; SEs a few percent apart (about 0.5% for the ATET, 4% for the potential-outcome mean on `cattaneo2`, with the default variance) |
| `teffects ..., pomeans` | `avg_predictions(fit, variables = "t")` | as the estimator's row |
| risk ratio after `teffects` (`nlcom` on the log scale) | `avg_comparisons(fit, variables = "t", comparison = "lnratioavg", transform = exp)` (recommended) | p-value against 1 and the exponentiated interval |
| `teffects ra (y x) (t)` | `lm(y ~ t * (x))` + `avg_comparisons(..., vcov = "HC0")` / `avg_predictions(..., vcov = "HC0")` | estimates; SEs about 0.1% apart (covariate sampling); 1.8% without `vcov = "HC0"` |
| `teffects ipwra` | `glm_weightit(y ~ t * (x), weightit = ...)` + marginaleffects | estimates; SEs 0.1-0.6% apart |
| `teffects aipw`, `psmatch`, `nnmatch` | a data frame of the estimates (`term`, `estimate`, `conf.low`, `conf.high`, `p.value`) | as supplied |
| multi-valued treatment, `control(#)` | a factor treatment, releveled to the control before fitting; `clean = TRUE` (a releveled factor's raw keys use its new positions, not Stata's codes) | estimates |
| `margins t` (factor `t`) | `avg_predictions(fit, variables = "t")` | estimates and SEs |
| `margins, dydx(x)` (continuous `x`) | `avg_slopes(fit, variables = "x")` | estimates and SEs |
| `margins, dydx(i.f)` | `avg_comparisons(fit, variables = "f")` (base level shown as `refcat`) | estimates and SEs |
| `margins, dydx(i.b)`, `b` a 0/1 variable | `avg_slopes(fit, variables = "b")` (the discrete change 1 - 0) | estimates and SEs |
| `margins, dydx(b)`, `b` 0/1 without `i.` | no marginaleffects counterpart (it always takes the discrete change of a 0/1 variable) | differ, up to about 9% (0.1-0.3% on the Stata demo cohort) |
| `margins, at(x = (...))` | `avg_predictions(fit, variables = list(x = c(...)))`: R shows the values under the variable label, Stata the raw `1._at` keys | estimates and SEs |
| `margins a, over(g)` | `avg_predictions(fit, variables = "a", by = c("g", "a"))`: R shows value labels, Stata the raw `1.g#0.a` keys | estimates and SEs |
| `margins` after `logit [pweight]` | `avg_*(..., vcov = tt_vcov(fit, "robust"), wts = w)` | estimates and SEs |
| `margins` after `logit [pweight], vce(cluster id)` (MSM) | `avg_*(..., vcov = tt_vcov(fit, "cluster", cluster = ~id), wts = w)` | estimates and SEs |
| `effecttab, from(M)` | the matrix: columns estimate, lower, upper, p; row names as labels | identical |

The agreement is with the output of Stata's `effecttab` on the same data.
Stata's `r(table)` of any of these commands, read into a data frame with its
`equation` and `term` (`r1vs0.t`, `0.t`, `1b.f`, `1._at`) columns, gives
Stata's table exactly, raw keys included.

## Part 2: a marginal structural model with time-varying treatment

When treatment varies over time and a time-varying risk factor both
predicts later treatment and is affected by earlier treatment, adjusting
for that factor in the outcome model removes part of the treatment effect,
and not adjusting leaves confounding. An MSM fitted with inverse
probability weights handles both (Robins, Hernán and Brumback 2000).

### Person-period data

1,000 people are followed for up to five periods. In each period a binary
risk factor `L` is measured, treatment `A` may start (and then continues),
and the outcome `Y` may occur, which ends follow-up. `L` raises the chance
of starting treatment and of the outcome, and treatment lowers later `L`.

```{r msm-data}
set.seed(20260927)
n_id <- 1000
n_per <- 5
V <- rbinom(n_id, 1, 0.4)                 # baseline risk factor
A_prev <- L_prev <- integer(n_id)
at_risk <- rep(TRUE, n_id)
periods <- vector("list", n_per)
for (k in seq_len(n_per)) {
  i <- which(at_risk)
  L <- rbinom(length(i), 1,
              plogis(-1 + 0.8 * L_prev[i] - 0.7 * A_prev[i] + 0.5 * V[i]))
  A <- ifelse(A_prev[i] == 1, 1L,
              rbinom(length(i), 1, plogis(-1.5 + 1.2 * L + 0.3 * V[i] + 0.1 * k)))
  Y <- rbinom(length(i), 1, plogis(-3.2 - 0.5 * A + 1.0 * L + 0.4 * V[i]))
  periods[[k]] <- data.frame(id = i, period = k, V = V[i], L = L,
                             A_prev = A_prev[i], A = A, Y = Y)
  A_prev[i] <- A
  L_prev[i] <- L
  at_risk[i[Y == 1]] <- FALSE
}
pp <- do.call(rbind, periods)
pp <- pp[order(pp$id, pp$period), ]
c(rows = nrow(pp), people = length(unique(pp$id)), events = sum(pp$Y))
```

### Stabilized time-varying weights

Treatment start is modelled among the rows still untreated at the end of
the previous period: the denominator model includes the time-varying `L`,
the numerator model only baseline `V` and period. Once treated, a person's
factor is 1. The weight in period k is the product of the factors up to k.

```{r msm-weights}
msm_weights <- function(pp) {
  untreated <- pp$A_prev == 0
  den <- glm(A ~ L + V + factor(period), family = binomial,
             data = pp[untreated, ])
  num <- glm(A ~ V + factor(period), family = binomial,
             data = pp[untreated, ])
  p_den <- predict(den, pp, type = "response")
  p_num <- predict(num, pp, type = "response")
  f <- ifelse(untreated,
              ifelse(pp$A == 1, p_num / p_den, (1 - p_num) / (1 - p_den)),
              1)
  ave(f, pp$id, FUN = cumprod)
}
pp$sw <- msm_weights(pp)
```

The ipw package (`ipw::ipwtm()`) builds the same kind of weights from the
two model formulas; `wttab()` reads its result's `$ipw.weights` directly
(tabtools does not depend on ipw).

Diagnostics by period, with `period =` (one row per period, as Cole and
Hernán 2008 tabulate them): the mean should stay close to 1, and a
drifting mean or a long right tail signals a misspecified weight model or
near-violations of positivity.

```{r msm-weight-summary}
wttab(pp, "sw", period = "period", digits = 3)
```

A common sensitivity analysis truncates the weights at the 1st and 99th
percentiles. `trunc =` adds the truncated weights beside the originals,
with the number of weights changed; the cut-offs are in `$stored$trunc`
(taken over all person-periods, as is usual for MSMs; `trunc_by =
"period"` takes them within each period):

```{r msm-truncate}
wt <- wttab(pp, "sw", trunc = c(0.01, 0.99), digits = 3)
wt
cut <- unlist(wt$stored$trunc[, c("lower", "upper")])
pp$sw_trunc <- pmin(pmax(pp$sw, cut[1]), cut[2])
```

### The weighted pooled logistic model

The MSM is a pooled logistic regression of the outcome on treatment and
baseline covariates, with period as a factor for the baseline hazard,
weighted by `sw`. Each person contributes several correlated rows, so the
variance is clustered by person: `vce = "cluster"` with `cluster = ~id`, as
Stata's `logit ... [pweight = sw], vce(cluster id)`.

```{r msm-fit}
pp$treated <- factor(pp$A, 0:1, c("Untreated", "Treated"))
pp$period_f <- factor(pp$period)
attr(pp$treated, "label") <- "Treatment"
attr(pp$period_f, "label") <- "Period"

crude_pl <- glm(Y ~ treated + period_f, family = binomial, data = pp)
cond_pl <- glm(Y ~ treated + V + L + period_f, family = binomial, data = pp)
msm <- suppressWarnings(glm(Y ~ treated + V + period_f, family = binomial,
                            data = pp, weights = sw))
msm_t <- suppressWarnings(glm(Y ~ treated + V + period_f, family = binomial,
                              data = pp, weights = sw_trunc))

t_msm <- regtab(crude_pl, cond_pl, msm, msm_t,
                models = c("Crude", "Adjusted for L", "MSM",
                           "MSM, truncated"),
                vce = "cluster", cluster = ~id, keep = "treated",
                stats = "n", vce_note = TRUE)
t_msm
```

`keep = "treated"` shows only the treatment rows, as MSM tables usually
do; `drop = "period_f"` would keep the covariates too. `Observations` counts
person-periods. The footnote records the clustered variance:

```{r msm-footnote}
t_msm$footnote
```

Truncation at the 1st and 99th percentiles changes nothing at two
decimals here; these weights have no extreme tail. The "Adjusted for L"
column conditions on a variable on the causal pathway
and is shown only for contrast; it is not an estimate of the marginal
effect.

The marginal effect of treatment on the per-period risk comes from the same
weighted model with marginaleffects, averaged with the weights and with the
clustered variance regtab used (Stata's `margins` after
`logit ... [pweight = sw], vce(cluster id)`), as a risk difference and, by
the log-ratio route, a risk ratio:

```{r msm-effecttab, eval = has("marginaleffects")}
V_msm <- tt_vcov(msm, "cluster", cluster = ~id)
effecttab(
  RD = marginaleffects::avg_comparisons(msm, variables = "treated",
                                        vcov = V_msm, wts = pp$sw),
  RR = marginaleffects::avg_comparisons(msm, variables = "treated",
                                        comparison = "lnratioavg",
                                        transform = exp,
                                        vcov = V_msm, wts = pp$sw),
  effect = "Per-period risk", digits = 3
)
```

### Other ways to fit the same MSM

The same weighted model can be fitted as a survey-weighted GLM with the
person as the sampling unit, which is how the ipw package's documentation
fits it, or as an independence GEE. regtab reads both: `svyglm()` with the
design-based variance and t on the design degrees of freedom (Stata's
`svy:`), and `geeglm()` with `gee_as = "glm"` (Stata's
`glm ... [pweight], vce(cluster id)`; by default a GEE is Stata's `xtgee`,
also shown as odds ratios, with xtgee's robust variance for a weighted fit).

```{r msm-alternatives, eval = has("survey", "geepack")}
design <- survey::svydesign(ids = ~id, weights = ~sw, data = pp)
msm_svy <- survey::svyglm(Y ~ treated + V + period_f, design = design,
                          family = quasibinomial)
msm_gee <- suppressWarnings(
  geepack::geeglm(Y ~ treated + V + period_f, family = binomial, data = pp,
                  id = id, weights = sw, corstr = "independence")
)
regtab(msm, msm_svy, msm_gee,
       models = c("glm, cluster", "svyglm", "geeglm"),
       vce = c("cluster", "stata", "stata"), cluster = list(~id, NULL, NULL),
       gee_as = "glm", keep = "treated")
```

`geeglm()` needs the data sorted by `id`, as they are here; regtab refuses
a fit on unsorted data, where geepack would split a person into several
clusters.

### A weighted Cox MSM

With counting-process data (`start`, `stop`) the MSM can be a weighted Cox
model. Give survival the person identifier with `id =` (or `cluster =`):
the robust variance then sums over each person's rows, and regtab applies
the G/(G - 1) factor.

```{r msm-cox}
pp$start <- pp$period - 1
pp$stop <- pp$period
msm_cox <- coxph(Surv(start, stop, Y) ~ treated + V, data = pp,
                 weights = sw, id = id, ties = "breslow")
t_cox <- regtab(msm_cox, models = "Cox MSM", keep = "treated",
                vce_note = TRUE)
t_cox
t_cox$footnote
```

### Bootstrap intervals

Bootstrap confidence intervals, resampling people and re-estimating the
weights in each sample, are often preferred for MSMs. regtab has no
bootstrap variance, but it tables supplied intervals through a data frame
(see [Regression tables](regression-tables.html#data-frame-input)). The
`key` column (`2.treated`, the second level of `treated`) puts the row on
the fitted model's treatment row. The loop below uses 100 replicates to keep
this article fast; use at least 1,000.

```{r msm-bootstrap}
set.seed(1)
rows_of <- split(seq_len(nrow(pp)), pp$id)
boot_log_or <- replicate(100, {
  pick <- sample(names(rows_of), replace = TRUE)
  b <- pp[unlist(rows_of[pick]), ]
  b$id <- rep(seq_along(pick), lengths(rows_of[pick]))
  b$sw <- msm_weights(b)
  fit <- suppressWarnings(glm(Y ~ treated + V + period_f, family = binomial,
                              data = b, weights = sw))
  coef(fit)[["treatedTreated"]]
})
boot_ci <- data.frame(
  term = "treatedTreated", key = "2.treated",
  estimate = exp(coef(msm)[["treatedTreated"]]),
  conf.low = exp(quantile(boot_log_or, 0.025)),
  conf.high = exp(quantile(boot_log_or, 0.975))
)
attr(boot_ci, "effect_scale") <- "OR"
attr(boot_ci, "conf.level") <- .95
regtab(msm, boot_ci, models = c("MSM, cluster-robust", "MSM, bootstrap"),
       vce = c("cluster", "stata"), cluster = list(~id, NULL),
       keep = "treated")
```

A data frame without a `p.value` column and with supplied intervals gets no
p-value: regtab derives one only from a standard error.

## What tabtools covers, and what it does not

| Output | tabtools |
|---|---|
| Baseline balance before and after weighting, SMDs, effective sample size | `table1_tc(wt =, wtcompare = TRUE, smd = TRUE)` |
| Balance in each period of a time-varying MSM | one `table1_tc()` per period, stacked into one sheet with `stacktab()` |
| Weight distribution and truncation table | `wttab()`: overall, by treatment (`by`), by period (`period`), truncation blocks (`trunc`), from a weight vector, a data frame, WeightIt or ipw |
| Weighted effect estimates with robust, cluster-robust, survey or M-estimation variance | `regtab()` with `vce`, `svyglm()`, `glm_weightit()`, weighted `coxph()`, `geeglm(gee_as = "glm")` |
| Crude, adjusted, IPTW and MSM estimates side by side | `regtab()` with one `vce` per model |
| Bootstrap intervals | a data frame with `conf.low`/`conf.high` |
| Treatment effects (ATE, ATET, potential-outcome means, risk ratios), predictive margins, marginal effects | `effecttab()` on marginaleffects results, a data frame or a matrix |
| Standardized survival curves | not a tabtools table; marginaleffects or adjustedCurves, then a data frame for `effecttab()` |
| Weighted models of classes without a robust variance (`polr`, `multinom`, `survreg`, mixed models) | refused for `vce = "robust"`/`"cluster"`; their weights act as frequency weights, and regtab warns when they are not integers |

The last row matters for IPTW: an inverse-probability weight passed to
`polr()`, `multinom()`, `zeroinfl()` or `survreg()` is treated by the fitting
function as a frequency weight, which inflates the sample size and
understates the standard errors. regtab warns about non-integer weights on
those fits and refuses `vce = "robust"` for them rather than print a
misleading interval; refit `survreg()` with `robust = TRUE` for a robust
AFT model.

The Stata equivalents of every call above (`[pweight]`, `vce(cluster id)`,
`svy:`, `teffects ipw`) are listed in
[Coming from Stata](coming-from-stata.html#weighted).

## References

Robins JM, Hernán MA, Brumback B (2000). Marginal structural models and
causal inference in epidemiology. *Epidemiology* 11(5), 550-560.

Cole SR, Hernán MA (2008). Constructing inverse probability weights for
marginal structural models. *American Journal of Epidemiology* 168(6),
656-664.

Austin PC, Stuart EA (2015). Moving towards best practice when using inverse
probability of treatment weighting (IPTW) using the propensity score to
estimate causal treatment effects in observational studies. *Statistics in
Medicine* 34(28), 3661-3679.
